---
title: "ggplot2 Themes"
author: 
  - name: Calvin J. Chiou
    email: jjchiou@nccu.edu.tw
    affiliations: 
    - name: National Chengchi University (NCCU)
      department: Department of Finance
date: "May 31, 2024"
format: 
  html: 
    toc: true

editor: visual
---

## Introduction

In this document, we will explore various themes and styles available in ggplot2.

```{r}
library(ggplot2)
library(ggthemes)

# Basic plot
p <- ggplot(mpg, aes(x=displ, y=hwy, color=class)) +
  geom_point() +
  labs(title="Default Theme")

p
```

Below are examples of applying different themes to your plot:

## Theme Minimal

```{r}
p + theme_minimal() +
  labs(title="Theme Minimal")
```

## Theme Classic

```{r}
p + theme_dark() +
  labs(title="Theme Dark")
```

## Theme Economist

```{r}
p + theme_economist() +
  labs(title="Theme Economist")
```

## Theme Economist - White

```{r}
p + theme_economist_white() +
  labs(title="Theme Economist - White")
```

## Theme Excel

```{r}
p + theme_excel() +
  labs(title="Theme Excel")
```

## Theme Excel - New

```{r}
p + theme_excel_new() +
  labs(title="Theme Excel - New")
```

## Theme Highcharts

```{r}
p + theme_hc() +
  labs(title="Theme Highcharts")
```

## Theme Google Docs

```{r}
p + theme_gdocs() +
  labs(title="Theme Google Docs")
```

## Theme 538

```{r}
p + theme_fivethirtyeight() +
  labs(title="Theme 538")
```

## Theme Stata

```{r}
p + theme_stata() +
  labs(title="Theme Stata")
```

## Theme Wall Street Journal (WSJ)

```{r}
p + theme_wsj() +
  labs(title="Theme Wall Street Journal")
```

## Theme Pander

```{r}
p + theme_pander() +
  labs(title="Theme Pander")
```

## Custom Theme

```{r}
custom_theme <- theme(
  plot.title = element_text(size=14, face="bold", hjust=0.5),
  axis.title = element_text(size=12, face="bold"),
  axis.text = element_text(size=10),
  panel.background = element_rect(fill="white"),
  panel.grid.major = element_line(color="grey", size=0.5),
  panel.grid.minor = element_line(color="lightgrey", size=0.25)
)

p + custom_theme + labs(title="Custom Theme")
```

## 
