---
title: "ggplot2 Theme Gallery"
author: 
  - name: "Asst. Prof. Calvin J. Chiou"
    email: jjchiou@nccu.edu.tw
    affiliations: 
    - name: "National Chengchi University (NCCU)"
format: 
  html: 
    toc: true
    toc-depth: 3
    toc-location: right
    number-sections: true
    self-contained: true
    highlight-style: github
    title-block-banner: true
editor: visual
---

```{=html}
<style>
  .quarto-title-block .quarto-title {
    display: flex;
    align-items: center;
    gap: 0.6em;
  }
  .quarto-title-block .quarto-title::before {
    content: "";
    display: inline-block;
    width: 48px;
    height: 48px;
    background: url("logo_ggplot2.png") no-repeat center/contain;
    flex-shrink: 0;
  }
</style>
```

# Setup

In this document, we explore various themes and styles available in `ggplot2` and related packages. All examples use a **line chart with three series** drawn from the built-in `economics_long` dataset (unemployment rate, personal consumption expenditure, and personal savings rate — rescaled to a common 0–100 index for comparability).

```{r setup, message=FALSE, warning=FALSE}
library(ggplot2)

# ── Base data: three economic series, re-indexed to 0–100 ──────────────────
rescale01 <- function(x) (x - min(x)) / (max(x) - min(x)) * 100

econ <- subset(economics_long,
               variable %in% c("uempmed", "pce", "psavert"))
econ$variable <- factor(econ$variable,
                        levels  = c("uempmed", "pce", "psavert"),
                        labels  = c("Unemployment", "Consumption", "Savings"))

# Index each series to [0, 100]
econ$value_idx <- ave(econ$value01, econ$variable,
                      FUN = function(x) rescale01(x))

# Base plot object (reused throughout)
p <- ggplot(econ, aes(x = date, y = value_idx, color = variable)) +
  geom_line(linewidth = 0.8) +
  labs(x = "Year", y = "Index (0 – 100)",
       color = "Series",
       caption = "Source: ggplot2::economics_long")
```

------------------------------------------------------------------------

# Built-in `ggplot2` Themes

Below are examples of every built-in theme shipped with `ggplot2`.

## Theme Gray (default)

```{r}
p + theme_gray() +
  labs(title = "Theme Gray")
```

## Theme Dark-on-light (BW)

```{r}
p + theme_bw() +
  labs(title = "Theme BW")
```

## Theme Linedraw

```{r}
p + theme_linedraw() +
  labs(title = "Theme Linedraw")
```

## Theme Light

```{r}
p + theme_light() +
  labs(title = "Theme Light")
```

## Theme Dark

```{r}
p + theme_dark() +
  labs(title = "Theme Dark")
```

## Theme Minimal

```{r}
p + theme_minimal() +
  labs(title = "Theme Minimal")
```

## Theme Classic

```{r}
p + theme_classic() +
  labs(title = "Theme Classic")
```

## Theme Void

```{r}
p + theme_void() +
  labs(title = "Theme Void")
```

## Theme Test

```{r}
p + theme_test() +
  labs(title = "Theme Test")
```

------------------------------------------------------------------------

# Themes in `ggthemes`

Additional themes and matching color scales from the `ggthemes` package (v 5.2.0, released 2025-11-30).

## Theme Base

Mimics base R graphics aesthetics. *New to the gallery — added in `ggthemes` 4.x.*

```{r}
library(ggthemes)

p + theme_base() +
  labs(title = "Theme Base")
```

## Theme Economist

```{r}
p + theme_economist() +
  scale_color_economist() +
  labs(title = "Theme Economist")
```

## Theme Economist – White

```{r}
p + theme_economist_white() +
  scale_color_economist() +
  labs(title = "Theme Economist – White")
```

## Theme Excel (Classic)

```{r}
p + theme_excel() +
  scale_color_excel() +
  labs(title = "Theme Excel (Classic)")
```

## Theme Excel – New

```{r, warning=FALSE}
p + theme_excel_new() +
  scale_color_excel_new() +
  labs(title = "Theme Excel – New")
```

## Theme LibreOffice Calc

```{r}
p + theme_calc() +
  scale_color_calc() +
  labs(title = "Theme Calc")
```

## Theme Highcharts

```{r}
p + theme_hc() +
  scale_color_hc() +
  labs(title = "Theme Highcharts")
```

## Theme Google Docs

```{r}
p + theme_gdocs() +
  scale_color_gdocs() +
  labs(title = "Theme Google Docs")
```

## Theme Clean

```{r, warning=FALSE}
p + theme_clean() +
  scale_color_canva() +
  labs(title = "Theme Clean")
```

## Theme 538

```{r}
p + theme_fivethirtyeight() +
  labs(title = "Theme 538")
```

## Theme igray

```{r}
p + theme_igray() +
  labs(title = "Theme igray")
```

## Theme Few

```{r}
p + theme_few() +
  scale_color_few() +
  labs(title = "Theme Few")
```

## Theme Solarized

```{r}
# Light
p + theme_solarized() +
  scale_color_solarized() +
  labs(title = "Theme Solarized – Light")

# Light 2
p + theme_solarized_2() +
  scale_color_solarized() +
  labs(title = "Theme Solarized 2")

# Dark
p + theme_solarized(light = FALSE) +
  scale_color_solarized() +
  labs(title = "Theme Solarized – Dark")
```

## Theme Tufte

```{r}
p + theme_tufte() +
  labs(title = "Theme Tufte")
```

## Theme Stata

```{r}
p + theme_stata() +
  scale_color_stata() +
  labs(title = "Theme Stata")
```

## Theme Wall Street Journal (WSJ)

```{r}
p + theme_wsj() +
  labs(title = "Theme Wall Street Journal")
```

## Theme Pander

```{r}
p + theme_pander() +
  scale_color_pander() +
  labs(title = "Theme Pander")
```

------------------------------------------------------------------------

## Notable Color Scales in `ggthemes`

The following demos use `theme_minimal()` as a neutral backdrop so the color scales are the focus.

### Colorblind-safe Palette

`scale_color_colorblind()` (alias: `scale_colour_colourblind()`) provides an 8-color palette that is safe for the most common forms of color-vision deficiency. *Renamed in `ggthemes` 5.2.0 to the British spelling.*

```{r}
p + theme_minimal() +
  scale_color_colorblind() +
  labs(title = "Colorblind-safe Palette (ggthemes)")
```

### Paul Tol Qualitative Palette

`scale_color_ptol()` implements Paul Tol's colorblind-safe qualitative palette, popular in scientific publishing.

```{r}
p + theme_minimal() +
  scale_color_ptol() +
  labs(title = "Paul Tol Qualitative Palette (ggthemes)")
```

### Tableau Color Scales

`scale_color_tableau()` offers Tableau's discrete palettes. The default is the classic "Tableau 10".

```{r}
p + theme_minimal() +
  scale_color_tableau() +
  labs(title = "Tableau 10 Palette (ggthemes)")
```

------------------------------------------------------------------------

# Themes in `tidyquant`

```{r, message=FALSE}
library(tidyquant)

# Tidyquant – standard
p + theme_tq() +
  scale_color_tq() +
  labs(title = "Theme Tidyquant")

# Tidyquant – dark
p + theme_tq_dark() +
  scale_color_tq() +
  labs(title = "Theme Tidyquant Dark")

# Tidyquant – green
p + theme_tq_green() +
  scale_color_tq() +
  labs(title = "Theme Tidyquant Green")
```

------------------------------------------------------------------------

# Themes in `hrbrthemes`

`hrbrthemes` (v 0.8.x, Bob Rudis) is a typography-focused package offering publication-ready themes. It ships with a bundled copy of Roboto Condensed so no font installation is required for the core themes.

```{r, message=FALSE, warning=FALSE}
library(hrbrthemes)

# Suppress font warnings if Roboto Condensed is unavailable on this system
update_geom_font_defaults()
```

## Theme ipsum

Clean, minimal, and opinionated — one of the most widely used non-ggplot2 themes in academic and data journalism contexts.

```{r, warning=FALSE}
p + theme_ipsum() +
  labs(title = "Theme ipsum (hrbrthemes)")
```

## Theme ipsum_rc (Roboto Condensed)

```{r, warning=FALSE}
p + theme_ipsum_rc() +
  labs(title = "Theme ipsum_rc (hrbrthemes)")
```

## Theme ft_rc (Financial Times – Dark)

Inspired by the Financial Times's dark-background online charts.

```{r, warning=FALSE}
p + theme_ft_rc() +
  labs(title = "Theme ft_rc – FT Dark (hrbrthemes)")
```

## Theme modern_rc (Modern Dark)

```{r, warning=FALSE}
p + theme_modern_rc() +
  labs(title = "Theme modern_rc – Modern Dark (hrbrthemes)")
```

------------------------------------------------------------------------

# Custom Theme

```{r}
custom_theme <- theme(
  plot.title       = element_text(size = 14, face = "bold", hjust = 0.5),
  axis.title       = element_text(size = 12, face = "bold"),
  axis.text        = element_text(size = 10),
  panel.background = element_rect(fill = "white"),
  panel.grid.major = element_line(color = "grey80",   linewidth = 0.5),
  panel.grid.minor = element_line(color = "grey90",   linewidth = 0.25),
  legend.position  = "bottom"
)

p + custom_theme +
  labs(title = "Custom Theme")
```

------------------------------------------------------------------------

# Academic Journal Style

Academic journals (e.g., *Journal of Finance*, *Review of Financial Studies*) typically require figures that are **black-and-white compatible**, use **grayscale or high-contrast linetype differentiation**, have **no chart junk**, and are sized for a **single or double column** layout. The example below builds a self-contained `theme_journal` that satisfies these conventions, and uses **linetypes instead of color** so the figure survives grayscale printing.

```{r journal-style, fig.width=5, fig.height=3.5, fig.dpi=300}
# ── Journal theme definition ───────────────────────────────────────────────
theme_journal <- function(base_size = 10, base_family = "serif") {
  theme_classic(base_size = base_size, base_family = base_family) %+replace%
    theme(
      # Title & caption
      plot.title      = element_text(size = base_size, face = "bold",
                                     hjust = 0, margin = margin(b = 4)),
      plot.caption    = element_text(size = base_size - 2, hjust = 0,
                                     color = "grey40",
                                     margin = margin(t = 6)),
      # Axes
      axis.title      = element_text(size = base_size),
      axis.text       = element_text(size = base_size - 1, color = "black"),
      axis.line       = element_line(color = "black", linewidth = 0.4),
      axis.ticks      = element_line(color = "black", linewidth = 0.4),
      # No background, no minor grid; faint major grid only
      panel.grid.major = element_line(color = "grey88", linewidth = 0.3),
      panel.grid.minor = element_blank(),
      # Legend: compact, inside or below
      legend.title    = element_text(size = base_size - 1, face = "bold"),
      legend.text     = element_text(size = base_size - 1),
      legend.key.width = unit(1.5, "cm"),   # room to see the linetype
      legend.position = "bottom",
      legend.background = element_blank(),
      # Margins
      plot.margin     = margin(6, 8, 6, 6)
    )
}

# ── Plot: grayscale + linetype (no reliance on color) ─────────────────────
p_journal <- ggplot(econ,
                    aes(x = date, y = value_idx,
                        color    = variable,
                        linetype = variable)) +
  geom_line(linewidth = 0.6) +
  scale_color_grey(start = 0, end = 0.6) +      # black → mid-grey
  scale_linetype_manual(values = c("solid", "dashed", "dotdash")) +
  labs(
    title    = "U.S. Economic Indicators, 1967–2015",
    subtitle = NULL,
    x        = "Year",
    y        = "Index (0 – 100)",
    color    = NULL,
    linetype = NULL,
    caption  = "Note: Each series re-indexed to [0, 100]. Source: ggplot2::economics_long."
  ) +
  theme_journal()

p_journal
```

Key design choices that match journal submission guidelines:

- **Serif font** (`"serif"`) matches the body typeface of most economics journals.
- **`scale_color_grey()` + `scale_linetype_manual()`** — the figure is fully legible in both color and grayscale print, satisfying most journals' figure requirements.
- **`fig.width = 5, fig.height = 3.5`** targets a single-column figure width (\~127 mm) common in two-column journal layouts.
- **`fig.dpi = 300`** meets the minimum resolution for most journals (300 dpi for line art).
- No fill, no background shading, no minor gridlines — eliminates chart junk per Tufte's principles.

To export the figure as a standalone file for submission:

```{r journal-export, eval=FALSE}
ggsave("fig1.png", plot = p_journal,
       width = 5, height = 3.5, units = "in", device = "png")

ggsave("fig1.tiff", plot = p_journal,
       width = 5, height = 3.5, units = "in",
       dpi = 600, compression = "lzw")   # 600 dpi TIFF for some journals
```
