---
title: "highcharter Themes"
author: 
  - name: "Asst. Prof. Calvin J. Chiou"
    email: jjchiou@nccu.edu.tw
    affiliations: 
    - name: "National Chengchi University (NCCU)"
format: 
  html: 
    toc: true
editor: visual
---

## Introduction

In this document, we will explore various themes and styles available in `highcharter`.

```{r, message=FALSE}
library(highcharter) 
library(dplyr)

# Sample dataset
data <- mtcars %>%
  tibble::rownames_to_column(var = "car") %>%
  mutate(cyl = as.factor(cyl))
# Basic plot
p <- highchart() %>% 
    hc_add_series(data, type = "scatter", hcaes(x = mpg, y = hp, group = cyl))
p %>% hc_title(text = "Default Theme")
```

Below are examples of applying different themes to your plot:

## Theme Google

```{r}
p %>%
  hc_title(text = "Theme Google") %>%
  hc_add_theme(hc_theme_google())
```

## Theme Financial Times

```{r}
p %>%
  hc_title(text = "Theme Financial Times") %>%
  hc_add_theme(hc_theme_ft())
```

## Theme Economist

```{r}
p %>%
  hc_title(text = "Theme Economist") %>%
  hc_add_theme(hc_theme_economist())
```

## Theme Elementary

```{r}
p %>%
  hc_title(text = "Theme Elementary") %>%
  hc_add_theme(hc_theme_elementary())
```

## Theme Simple

```{r}
p %>%
  hc_title(text = "Theme Simple") %>%
  hc_add_theme(hc_theme_smpl())
```

## Theme Database

```{r}
p %>%
  hc_title(text = "Theme Database") %>%
  hc_add_theme(hc_theme_db())
```

## Theme FFX

```{r}
p %>%
  hc_title(text = "Theme FFX") %>%
  hc_add_theme(hc_theme_ffx())
```

## Theme FiveThirtyEight

```{r}
p %>%
  hc_title(text = "Theme FiveThirtyEight") %>%
  hc_add_theme(hc_theme_538())
```

## Theme Flat

```{r}
p %>%
  hc_title(text = "Theme Flat") %>%
  hc_add_theme(hc_theme_flat())
```

## Theme Flat Dark

```{r}
p %>%
  hc_title(text = "Theme Flat Dark") %>%
  hc_add_theme(hc_theme_flatdark())
```

## Theme Chalk

```{r}
p %>%
  hc_title(text = "Theme Chalk") %>%
  hc_add_theme(hc_theme_chalk())
```

## Theme Handdrawn

```{r}
p %>%
  hc_title(text = "Theme Handdrawn") %>%
  hc_add_theme(hc_theme_handdrawn())
```

## Theme Dark Unica

```{r}
p %>%
  hc_title(text = "Theme Dark Unica") %>%
  hc_add_theme(hc_theme_darkunica())
```

## Theme ggplot2

```{r}
p %>%
  hc_title(text = "Theme ggplot2") %>%
  hc_add_theme(hc_theme_ggplot2())
```

## Theme Alone

```{r}
p %>%
  hc_title(text = "Theme Alone") %>%
  hc_add_theme(hc_theme_alone())
```

## Theme Bloom

```{r}
p %>%
  hc_title(text = "Theme Bloom") %>%
  hc_add_theme(hc_theme_bloom())
```

## Theme Gridlight

```{r}
p %>%
  hc_title(text = "Theme Grid Light") %>%
  hc_add_theme(hc_theme_gridlight())
```

## Theme Highcharter

```{r}
p %>%
  hc_title(text = "Theme Highcharter") %>%
  hc_add_theme(hc_theme_hcrt())
```

## Theme Null

```{r}
p %>%
  hc_title(text = "Theme Null") %>%
  hc_add_theme(hc_theme_null())
```

## Theme Monokai

```{r}
p %>%
  hc_title(text = "Theme Monokai") %>%
  hc_add_theme(hc_theme_monokai())
```

## Theme Sansignika

```{r}
p %>%
  hc_title(text = "Theme Sansignika") %>%
  hc_add_theme(hc_theme_sandsignika())
```

## Theme Sparkline

```{r}
p %>%
  hc_title(text = "Theme Sparkline") %>%
  hc_add_theme(hc_theme_sparkline())
p %>%
  hc_title(text = "Theme Sparkline VB") %>%
  hc_add_theme(hc_theme_sparkline_vb())
```

## Theme Superheroes

```{r}
p %>%
  hc_title(text = "Theme Superheroes") %>%
  hc_add_theme(hc_theme_superheroes())
```

## Theme Tufte

```{r}
p %>%
  hc_title(text = "Theme Tufte") %>%
  hc_add_theme(hc_theme_tufte())
p %>%
  hc_title(text = "Theme Tufte2") %>%
  hc_add_theme(hc_theme_tufte2())
```
