---
title: "highcharter Theme Gallery"
author: 
  - name: "Asst. Prof. Calvin J. Chiou"
    email: jjchiou@nccu.edu.tw
    affiliations: 
    - name: "National Chengchi University (NCCU)"
format: 
  html: 
    toc: true
    toc-depth: 3
    toc-location: right
    number-sections: true
    self-contained: true
editor: visual
---

```{=html}
<style>
  .quarto-title-block .quarto-title {
    display: flex;
    align-items: center;
    gap: 0.6em;
  }
  .quarto-title-block .quarto-title::before {
    content: "";
    display: inline-block;
    width: 228px;
    height: 48px;
    background: url("highcharts-logo.png") no-repeat center/contain;
    flex-shrink: 0;
  }
</style>
```

# Setup

In this document, we explore various themes available in `highcharter` — the R wrapper for the Highcharts JavaScript library. All examples use a **line chart with three series** drawn from the built-in `economics_long` dataset (unemployment duration, personal consumption expenditure, and personal savings rate — rescaled to a common 0–100 index for comparability), mirroring the dataset used in the companion `ggplot2` theme gallery.

```{r setup, message=FALSE, warning=FALSE}
library(highcharter)
library(dplyr)

# ── Base data: three economic series, re-indexed to 0–100 ──────────────────
rescale01 <- function(x) (x - min(x)) / (max(x) - min(x)) * 100

econ <- ggplot2::economics_long |>
  filter(variable %in% c("uempmed", "pce", "psavert")) |>
  mutate(
    variable = factor(variable,
                      levels = c("uempmed", "pce", "psavert"),
                      labels = c("Unemployment", "Consumption", "Savings")),
    year     = as.integer(format(date, "%Y"))
  ) |>
  group_by(variable) |>
  mutate(value_idx = rescale01(value01)) |>
  ungroup()

# ── Base highchart object (reused throughout) ──────────────────────────────
# Each series is added separately so highcharter draws three distinct lines.
make_base <- function() {
  unemp <- filter(econ, variable == "Unemployment")
  cons  <- filter(econ, variable == "Consumption")
  sav   <- filter(econ, variable == "Savings")

  highchart(type = "stock") |>
    hc_add_series(unemp, "line", hcaes(x = date, y = value_idx),
                  name = "Unemployment") |>
    hc_add_series(cons,  "line", hcaes(x = date, y = value_idx),
                  name = "Consumption") |>
    hc_add_series(sav,   "line", hcaes(x = date, y = value_idx),
                  name = "Savings") |>
    hc_xAxis(title = list(text = "Year")) |>
    hc_yAxis(title = list(text = "Index (0 – 100)")) |>
    hc_tooltip(shared = TRUE, valueDecimals = 1) |>
    hc_credits(enabled = TRUE,
               text   = "Source: ggplot2::economics_long")
}
```

> **Note:** `highcharter` produces **interactive HTML widgets**. Hover over any line to see crosshair tooltips; click legend entries to toggle series visibility. This is one of `highcharter`'s key advantages over static `ggplot2` output.

------------------------------------------------------------------------

# Default Theme

```{r}
make_base() |>
  hc_title(text = "Default Theme")
```

------------------------------------------------------------------------

# Built-in `highcharter` Themes

All themes below are applied with `hc_add_theme()`. Each theme ships as a named function `hc_theme_*()` that returns an `hc_theme` object.

## Theme Google

```{r}
make_base() |>
  hc_title(text = "Theme Google") |>
  hc_add_theme(hc_theme_google())
```

## Theme Financial Times

```{r}
make_base() |>
  hc_title(text = "Theme Financial Times") |>
  hc_add_theme(hc_theme_ft())
```

## Theme Economist

```{r}
make_base() |>
  hc_title(text = "Theme Economist") |>
  hc_add_theme(hc_theme_economist())
```

## Theme Elementary

```{r}
make_base() |>
  hc_title(text = "Theme Elementary") |>
  hc_add_theme(hc_theme_elementary())
```

## Theme Simple

```{r}
make_base() |>
  hc_title(text = "Theme Simple") |>
  hc_add_theme(hc_theme_smpl())
```

## Theme Database

```{r}
make_base() |>
  hc_title(text = "Theme Database") |>
  hc_add_theme(hc_theme_db())
```

## Theme FFX

```{r}
make_base() |>
  hc_title(text = "Theme FFX") |>
  hc_add_theme(hc_theme_ffx())
```

## Theme FiveThirtyEight

```{r}
make_base() |>
  hc_title(text = "Theme FiveThirtyEight") |>
  hc_add_theme(hc_theme_538())
```

## Theme Flat

```{r}
make_base() |>
  hc_title(text = "Theme Flat") |>
  hc_add_theme(hc_theme_flat())
```

## Theme Flat Dark

```{r}
make_base() |>
  hc_title(text = "Theme Flat Dark") |>
  hc_add_theme(hc_theme_flatdark())
```

## Theme Chalk

```{r}
make_base() |>
  hc_title(text = "Theme Chalk") |>
  hc_add_theme(hc_theme_chalk())
```

## Theme Handdrawn

```{r}
make_base() |>
  hc_title(text = "Theme Handdrawn") |>
  hc_add_theme(hc_theme_handdrawn())
```

## Theme Dark Unica

```{r}
make_base() |>
  hc_title(text = "Theme Dark Unica") |>
  hc_add_theme(hc_theme_darkunica())
```

## Theme ggplot2

Replicates the classic `ggplot2` gray aesthetic inside Highcharts.

```{r}
make_base() |>
  hc_title(text = "Theme ggplot2") |>
  hc_add_theme(hc_theme_ggplot2())
```

## Theme Alone

```{r}
make_base() |>
  hc_title(text = "Theme Alone") |>
  hc_add_theme(hc_theme_alone())
```

## Theme Bloom

```{r}
make_base() |>
  hc_title(text = "Theme Bloom") |>
  hc_add_theme(hc_theme_bloom())
```

## Theme Grid Light

```{r}
make_base() |>
  hc_title(text = "Theme Grid Light") |>
  hc_add_theme(hc_theme_gridlight())
```

## Theme Highcharter

The default documentation theme for the `highcharter` package itself.

```{r}
make_base() |>
  hc_title(text = "Theme Highcharter") |>
  hc_add_theme(hc_theme_hcrt())
```

## Theme Null

Strips all styling — useful as a blank canvas for fully custom themes.

```{r}
make_base() |>
  hc_title(text = "Theme Null") |>
  hc_add_theme(hc_theme_null())
```

## Theme Monokai

```{r}
make_base() |>
  hc_title(text = "Theme Monokai") |>
  hc_add_theme(hc_theme_monokai())
```

## Theme Sandsignika

```{r}
make_base() |>
  hc_title(text = "Theme Sandsignika") |>
  hc_add_theme(hc_theme_sandsignika())
```

## Theme Sparkline

`hc_theme_sparkline()` and its variant `hc_theme_sparkline_vb()` strip axes and chrome to produce compact inline sparklines.

```{r}
make_base() |>
  hc_title(text = "Theme Sparkline") |>
  hc_add_theme(hc_theme_sparkline())

make_base() |>
  hc_title(text = "Theme Sparkline VB") |>
  hc_add_theme(hc_theme_sparkline_vb())
```

## Theme Superheroes

```{r}
make_base() |>
  hc_title(text = "Theme Superheroes") |>
  hc_add_theme(hc_theme_superheroes())
```

## Theme Tufte

Inspired by Edward Tufte's data-ink ratio principles. `hc_theme_tufte2()` is a lighter variant.

```{r}
make_base() |>
  hc_title(text = "Theme Tufte") |>
  hc_add_theme(hc_theme_tufte())

make_base() |>
  hc_title(text = "Theme Tufte 2") |>
  hc_add_theme(hc_theme_tufte2())
```

------------------------------------------------------------------------

# Custom Theme

`hc_theme()` constructs a theme from scratch, and `hc_theme_merge()` layers a custom override on top of any existing theme — analogous to `%+replace%` in `ggplot2`.

```{r}
# ── Stand-alone custom theme ───────────────────────────────────────────────
custom_hc_theme <- hc_theme(
  colors = c("#2C3E50", "#E74C3C", "#27AE60"),   # 3 series colors
  chart  = list(
    backgroundColor = "#FAFAFA",
    style = list(fontFamily = "Georgia, serif")
  ),
  title = list(
    style = list(color = "#2C3E50", fontSize = "16px", fontWeight = "bold")
  ),
  xAxis = list(
    gridLineColor  = "#ECF0F1",
    lineColor      = "#BDC3C7",
    tickColor      = "#BDC3C7",
    labels = list(style = list(color = "#7F8C8D"))
  ),
  yAxis = list(
    gridLineColor  = "#ECF0F1",
    labels = list(style = list(color = "#7F8C8D"))
  ),
  legend = list(
    itemStyle      = list(color = "#2C3E50", fontWeight = "normal"),
    itemHoverStyle = list(color = "#E74C3C")
  ),
  tooltip = list(
    backgroundColor = "#2C3E50",
    style           = list(color = "#FFFFFF")
  )
)

make_base() |>
  hc_title(text = "Custom Theme") |>
  hc_add_theme(custom_hc_theme)
```

## Merging Themes with `hc_theme_merge()`

`hc_theme_merge()` takes any existing theme and patches only the properties you specify, leaving the rest intact — useful for minor tweaks without rewriting a full theme.

```{r}
# Start from Theme Simple and override just the background and title color
merged_theme <- hc_theme_merge(
  hc_theme_smpl(),
  hc_theme(
    chart = list(backgroundColor = "#F0F4F8"),
    title = list(style = list(color = "#1A237E", fontWeight = "bold"))
  )
)

make_base() |>
  hc_title(text = "Theme Simple + Custom Override") |>
  hc_add_theme(merged_theme)
```

------------------------------------------------------------------------

# Academic Journal Style

For working papers, slide decks, and online appendices, a clean high-contrast interactive figure is preferable to a static image. The theme below follows the same principles as the `theme_journal` in the `ggplot2` gallery — minimal chrome, serif font, no decorative fills — while retaining `highcharter`'s interactive tooltip and legend-toggle features.

```{r}
# ── Journal-style highcharter theme ───────────────────────────────────────
# ── Convert dates to JS timestamps (required for type = "chart" datetime axis)
econ_ts <- econ |> 
  mutate(ts = datetime_to_timestamp(date))

unemp <- filter(econ_ts, variable == "Unemployment")
cons  <- filter(econ_ts, variable == "Consumption")
sav   <- filter(econ_ts, variable == "Savings")

highchart() |>
  hc_chart(type = "line") |>
  hc_xAxis(type  = "datetime",
           title = list(text = "Year"),
           lineColor = "#000000", lineWidth = 1,
           tickColor = "#000000", gridLineWidth = 0,
           labels = list(style = list(color = "#000000", fontSize = "11px"))) |>
  hc_yAxis(title = list(text = "Index (0 – 100)"),
           gridLineColor = "#DDDDDD", gridLineWidth = 1,
           labels = list(style = list(color = "#000000", fontSize = "11px"))) |>
  hc_add_series(name = "Unemployment",
                data = list_parse2(data.frame(unemp$ts, unemp$value_idx))) |>
  hc_add_series(name = "Consumption",
                data = list_parse2(data.frame(cons$ts,  cons$value_idx)))  |>
  hc_add_series(name = "Savings",
                data = list_parse2(data.frame(sav$ts,   sav$value_idx)))   |>
  hc_colors(c("#000000", "#555555", "#AAAAAA")) |>      # explicit, not via theme
  hc_plotOptions(series = list(lineWidth = 1.5,
                               marker    = list(enabled = FALSE))) |>
  hc_tooltip(shared = TRUE, valueDecimals = 1,
             backgroundColor = "#FFFFFF", borderColor = "#000000",
             style = list(color = "#000000", fontSize = "11px")) |>
  hc_title(text = "U.S. Economic Indicators, 1967–2015",
           align = "left",
           style = list(fontSize = "13px", fontWeight = "bold")) |>
  hc_subtitle(text = "Each series re-indexed to [0, 100]",
              align = "left",
              style = list(color = "#444444", fontSize = "11px")) |>
  hc_legend(itemStyle      = list(color = "#000000", fontWeight = "normal",
                                  fontSize = "11px"),
            itemHoverStyle = list(color = "#555555")) |>
  hc_credits(enabled = TRUE,
             text  = "Source: ggplot2::economics_long",
             style = list(color = "#888888", fontSize = "10px")) |>
  hc_chart(style = list(fontFamily = "Georgia, Times New Roman, serif"))
```

Key design choices:

- **Serif font** (`Georgia`) matches the body typeface of most economics journals and is consistent with the `ggplot2` journal theme in the companion gallery.
- **Grayscale color palette** — black, mid-grey, light-grey — ensures the chart survives greyscale printing if a static screenshot is embedded in a PDF submission.
- **No dot markers** on lines (`marker = list(enabled = FALSE)`) reduces visual clutter in dense time-series charts.
- **Left-aligned title** follows APA and most economics journal figure conventions.
- **Minimal gridlines** — horizontal only, light grey — provides y-axis orientation without chart junk.
- Interactive tooltips and legend toggles are retained for the HTML/online-appendix version.
