---
title: "R Note 1 - Introduction to R and Financial Data Handling"
author: 
  - name: "Asst. Prof. Calvin J. Chiou"
    email: jjchiou@nccu.edu.tw
    affiliations: 
    - name: "National Chengchi University (NCCU)"
format: html
editor: visual
---

## Course Information

-   **Course Name:** Investments \| 000351051

-   **Semester:** Spring 2025

# Introduction to R & RStudio

Welcome to the first session of Investments! This tutorial introduces the **R programming language** and how to handle **financial data** for investment research.

## 1. Setting Up R and RStudio

Before we begin, ensure you have installed: - [R](https://www.r-project.org/) - [RStudio](https://posit.co/downloads/)

```{r, eval=FALSE}
# Install packages
install.packages("tidyquant")
install.packages("tidyverse")

```

```{r, message=FALSE}
# Load packages
library(tidyquant)
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(highcharter)
library(DT)
```

## 2. Basic R Syntax

```{r}
# Basic operations in R
x <- 10
y <- 5

sum_xy <- x + y
diff_xy <- x - y
prod_xy <- x * y
quot_xy <- x / y
exp_xy <- x^y

print(sum_xy)
print(diff_xy)
print(prod_xy)
print(quot_xy)
print(exp_xy)
```

## 3. Understanding Data Types in R

### Numeric

```{r}
num_var <- 42
print(num_var)
```

### Character (String)

```{r}
char_var <- "Hello, R!"
print(char_var)
```

### Logical (Boolean)

```{r}
bool_var <- TRUE
print(bool_var)
```

### Factor (Categorical Data)

```{r}
factor_var <- factor(c("low", "medium", "high", "medium"))
print(factor_var)
```

### Data Structure

```{r}
# Vectors
prices <- c(100, 102, 105, 107, 110)
print(prices)

# Lists (Collection of different types)
list_var <- list(num_var, char_var, bool_var, prices)
print(list_var)

# Matrices (2D array)
matrix_var <- matrix(1:9, nrow=3)
print(matrix_var)

# Data Frames
stock_data <- data.frame(
  Date = as.Date('2024-01-01') + 0:4,
  Price = prices
)
print(stock_data)
```

## 4. The Pipe Operator (`%>%` and `|>`)

The **pipe operator** allows you to pass the output of one function directly into another function, making your code cleaner and easier to read.

### Using `%>%` from `dplyr` (or `tidyverse` family)

The `dplyr` package provides the `%>%` pipe operator, which allows for chaining multiple operations.

```{r}
# Example dataset
data <- data.frame(x = 1:5, y = c(2, 4, 6, 8, 10))

data %>% 
  mutate(z = x + y) %>% 
  filter(z > 6) %>% 
  arrange(desc(z))
```

### Using `|>` (Base R Pipe)

R introduced the native `|>` pipe operator in version 4.1.0. It works similarly but has slightly different behavior:

```{r}
data |> 
  mutate(z = x + y) |> 
  filter(z > 6) |> 
  arrange(desc(z))
```

**Key Differences:**

-   `%>%` allows passing arguments to functions with `.` placeholder (useful for non-standard functions).

-   `|>` is optimized for performance and is recommended for base R users.

```{r}
# An alternative example:
# normal use of a function
mean(prices)
# pipe from dplyr
prices %>% mean
# pipe in base R
prices |> mean()
```

## 5. Using For Loops in R

A `for` loop allows us to **iterate** over a sequence and perform operations repeatedly.

### Basic For Loop Example

```{r}
# Print numbers from 1 to 5
for (i in 1:5) {
  print(i)
}
```

### Looping Over a Vector

```{r}
prices <- c(100, 102, 105, 107, 110)

for (price in prices) {
  print(price * 1.1)  # Increase price by 10%
}
```

### Storing Loop Output in a Vector

```{r}
adjusted_prices <- numeric(length(prices))

for (i in seq_along(prices)) {
  adjusted_prices[i] <- prices[i] * 1.1
}
print(adjusted_prices)
```

### Nesting Loops Example

```{r}
for (i in 1:3) {
  for (j in 1:2) {
    print(paste("i:", i, "j:", j))
  }
}
```

## 6. Importing Financial Data

Using `tidyquant` to fetch stock prices:

```{r}
# Get Apple (AAPL) stock data
apple_stock <- tq_get("AAPL", from = "2023-01-01", to = "2024-01-01")
# Interactive table
datatable(apple_stock) |> 
  formatRound(columns = c("open","high","low","close","adjusted"), digits = 2)
```

## 7. Basic Data Manipulation with `dplyr`

```{r}
# Calculate daily returns
apple_stock <- apple_stock |> 
  arrange(date) |> 
  mutate(daily_return = adjusted / lag(adjusted) - 1)
head(apple_stock)
```

## 8. Visualizing Stock Prices with `ggplot2`

The `ggplot2` is a powerful tool for creating visualizations in R. It follows the **Grammar of Graphics** approach, allowing users to build complex charts by adding multiple layers.

### Key Features of `ggplot2`:

-   Layered approach to building plots.
-   Highly customizable aesthetic mappings.
-   Supports multiple themes for different styling.

### Using Themes with `ggthemes`

The `ggthemes` package provides multiple predefined themes to enhance the aesthetics of plots. More styles of ggthemes can be found via [*ggplot2 Themes and Styles*](/notes/ggthemes.qmd)*.*

Now, let's plot the stock price trend for **AAPL** and use the wall street journal (WSJ) theme:

```{r}
# Plot price trend
ggplot(apple_stock, aes(x = date, y = adjusted)) +
  geom_line(color = "blue", linewidth = 1) +
  labs(title = "AAPL Stock Price", x = "Date", y = "Adjusted Close") +
  theme_wsj()
```

## 9. Interactive Visualization with `highcharter`

More pre-built themes of hicharter can be found via [*highcharter Themes and Styles*](/notes/hcthemes.qmd)*.*

```{r}
# Create interactive stock price and volume plot
highchart(type = "stock") |> 
  hc_add_series(apple_stock, type = "line", hcaes(x = date, y = adjusted), name = "AAPL Price") |> 
  hc_add_series(apple_stock, type = "column", hcaes(x = date, y = volume), name = "Trading Volume", yAxis = 1) |> 
  hc_yAxis_multiples(
    list(title = list(text = "Stock Price")),
    list(title = list(text = "Trading Volume"), opposite = TRUE)
  ) |> 
  hc_title(text = "AAPL Stock Price and Trading Volume") |>
  hc_add_theme(hc_theme_ft())

# Candlestick Chart
highchart(type = "stock") |> 
  hc_add_series(apple_stock, type = "candlestick", hcaes(x = date, open = open, high = high, low = low, close = adjusted), name = "AAPL Price") |> 
  hc_title(text = "AAPL Candlestick Chart")
```

## Exercises

1.  Fetch stock price data for **MSFT** using `tidyquant`.
2.  Compute daily returns for MSFT.
3.  Create a price trend plot for MSFT.
4.  Create a **vector, matrix, and data frame**, and perform calculations on them.
5.  Use **`highcharter`** to plot **MSFT price and trading volume**.

## Resources

1.  [Cookbook for R](http://www.cookbook-r.com/).

2.  [R Workflow](https://hbiostat.org/rflow/).

3.  GenAI: ChatGPT, Gemini, etc.

4.  [Data Analysis with R](https://www.coursera.org/specializations/statistics), Coursera, [Mine Çetinkaya-Rundel](https://www.coursera.org/instructor/minecetinkayarundel), Duke University.

5.  [Google Data Analytics](https://www.coursera.org/professional-certificates/google-data-analytics), Coursera, Google.

6.  [Stack Overflow](https://stackoverflow.com/questions) for coding Q&A!

------------------------------------------------------------------------

Happy coding! 🚀
