---
title: "R Note 4 - Trading Strategy"
author: 
  - name: "Asst. Prof. Calvin J. Chiou"
    email: jjchiou@nccu.edu.tw
    affiliations: 
    - name: "National Chengchi University (NCCU)"

format: 
  html: 
    toc: true
    toc-depth: 3
    toc-location: right
    number-sections: false
    self-contained: true
    highlight-style: github
editor: visual
---

```{r, include=FALSE}
library(tidyquant)
library(dplyr)
library(lubridate)
library(highcharter)
library(TTR)
```

## Course Information

- **Course Name:** Investments \| 000351051

- **Semester:** Spring 2026

# Introduction

This document demonstrates several quantitative strategies applied to the Taiwan 50 ETF (symbol: `0050.TW`), one of the most actively traded ETFs representing large-cap stocks in Taiwan. We cover four approaches:

1.  **Moving Average (MA) Crossover** — a trend-following strategy based on short- vs. long-term moving averages.
2.  **RSI Strategy** — a mean-reversion strategy using the Relative Strength Index.
3.  **Bollinger Bands Strategy** — a volatility-based mean-reversion strategy.
4.  **Dollar-Cost Averaging (DCA)** — a passive strategy that invests a fixed amount of NT\$20,000 each month, regardless of market conditions.

All strategies start with the same initial capital of NT\$1,000,000 over the period **January 1, 2023 to January 1, 2026**. A buy-and-hold 0050.TW position serves as the common benchmark throughout.

------------------------------------------------------------------------

## 1. Retrieve Historical Price Data

We begin by retrieving daily adjusted price data for 0050.TW. This single dataset drives all four strategies.

```{r}
etf_data <- tq_get("0050.TW",
                   from = "2023-01-01",
                   to   = "2026-01-01")
head(etf_data)
```

------------------------------------------------------------------------

## 2. Moving Average (MA) Crossover Strategy

The MA crossover is the baseline strategy. We use a short-term MA20 and a long-term MA50 to generate buy/sell signals.

### **2.1. Calculate Moving Averages**

```{r}
etf_ma <- etf_data |> 
  mutate(
    MA20 = SMA(adjusted, n = 20),
    MA50 = SMA(adjusted, n = 50)
  )
head(etf_ma)
```

### **2.2. Generate Trading Signals**

- **Buy** when MA20 crosses **above** MA50 (golden cross)
- **Sell** when MA20 crosses **below** MA50 (death cross)

```{r}
etf_ma_signals <- etf_ma |> 
  mutate(
    signal = case_when(
      lag(MA20) < lag(MA50) & MA20 >= MA50 ~ "buy",
      lag(MA20) > lag(MA50) & MA20 <= MA50 ~ "sell",
      TRUE ~ NA_character_
    )
  )

etf_ma_signals |> 
  filter(!is.na(signal)) |> 
  select(date, adjusted, MA20, MA50, signal)
```

### **2.3. Simulate the MA Strategy**

```{r}
capital  <- 1000000
position <- 0
cash     <- capital
portfolio_ma <- data.frame()

for (i in seq_len(nrow(etf_ma_signals))) {
  date_i  <- etf_ma_signals$date[i]
  price_i <- etf_ma_signals$adjusted[i]
  sig     <- etf_ma_signals$signal[i]

  if (!is.na(sig)) {
    if (sig == "buy") {
      position_new <- floor(cash / price_i)
      position <- position + position_new
      cash     <- cash - position_new * price_i
    } else if (sig == "sell") {
      cash     <- cash + position * price_i
      position <- 0
    }
  }
  total_value <- cash + position * price_i
  portfolio_ma <- bind_rows(portfolio_ma,
                   data.frame(date = date_i, price = price_i,
                              position, cash, total_value))
}

final_ma  <- tail(portfolio_ma$total_value, 1)
return_ma <- (final_ma - capital) / capital * 100
cat("MA Strategy — Final Value: NT$", round(final_ma, 0),
    "| Total Return:", round(return_ma, 2), "%\n")
```

### **2.4. Visualize MA Indicators and Signals**

```{r}
#| code-fold: true
# Mark buy/sell dates for flag annotation
buy_signals  <- etf_ma_signals |> filter(signal == "buy")
sell_signals <- etf_ma_signals |> filter(signal == "sell")

highchart(type = "stock") |>
  hc_title(text = "0050.TW — MA Crossover Signals") |>
  hc_add_series(etf_ma, type = "line", hcaes(x = date, y = adjusted),
                name = "0050 Price", color = "#234E70") |>
  hc_add_series(etf_ma, type = "line", hcaes(x = date, y = MA20),
                name = "MA20", color = "#E07A5F", dashStyle = "ShortDash") |>
  hc_add_series(etf_ma, type = "line", hcaes(x = date, y = MA50),
                name = "MA50", color = "#3D405B", dashStyle = "ShortDot") |>
  hc_add_series(buy_signals,  type = "scatter", hcaes(x = date, y = adjusted),
                name = "Buy",  color = "green",  marker = list(symbol = "triangle",   size = 8)) |>
  hc_add_series(sell_signals, type = "scatter", hcaes(x = date, y = adjusted),
                name = "Sell", color = "red",    marker = list(symbol = "triangle-down", size = 8)) |>
  hc_tooltip(valueDecimals = 2, shared = FALSE) |>
  hc_legend(enabled = T, align = "center", verticalAlign = "bottom") |>
  hc_add_theme(hc_theme_smpl())
```

------------------------------------------------------------------------

## 3. RSI Strategy

The **Relative Strength Index (RSI)** measures the speed and change of recent price movements on a scale from 0 to 100. Unlike the MA crossover, which is a trend-following rule, RSI is a **mean-reversion** indicator: extreme readings suggest the asset is overbought or oversold and likely to reverse.

$$RSI = 100 - \frac{100}{1 + RS}, \quad RS = \frac{\text{Average Gain (14 days)}}{\text{Average Loss (14 days)}}$$

The standard thresholds are:

- **RSI \< 30** → oversold → **Buy**
- **RSI \> 70** → overbought → **Sell**

### **3.1. Compute RSI**

```{r}
etf_rsi <- etf_data |> 
  mutate(RSI = RSI(adjusted, n = 14))
head(etf_rsi)
```

### **3.2. Generate RSI Signals**

We enter a position when RSI drops below 30 and exit when it rises above 70.

```{r}
etf_rsi_signals <- etf_rsi |> 
  mutate(
    signal = case_when(
      lag(RSI) >= 30 & RSI < 30  ~ "buy",   # crosses into oversold
      lag(RSI) <= 70 & RSI > 70  ~ "sell",  # crosses into overbought
      TRUE ~ NA_character_
    )
  )

etf_rsi_signals |> 
  filter(!is.na(signal)) |> 
  select(date, adjusted, RSI, signal)
```

### **3.3. Simulate the RSI Strategy**

```{r}
position <- 0
cash     <- capital
portfolio_rsi <- data.frame()

for (i in seq_len(nrow(etf_rsi_signals))) {
  date_i  <- etf_rsi_signals$date[i]
  price_i <- etf_rsi_signals$adjusted[i]
  sig     <- etf_rsi_signals$signal[i]

  if (!is.na(sig)) {
    if (sig == "buy") {
      position_new <- floor(cash / price_i)
      position <- position + position_new
      cash     <- cash - position_new * price_i
    } else if (sig == "sell") {
      cash     <- cash + position * price_i
      position <- 0
    }
  }
  total_value <- cash + position * price_i
  portfolio_rsi <- bind_rows(portfolio_rsi,
                    data.frame(date = date_i, price = price_i,
                               position, cash, total_value))
}

final_rsi  <- tail(portfolio_rsi$total_value, 1)
return_rsi <- (final_rsi - capital) / capital * 100
cat("RSI Strategy — Final Value: NT$", round(final_rsi, 0),
    "| Total Return:", round(return_rsi, 2), "%\n")
```

### **3.4. Visualize RSI and Signals**

```{r}
#| code-fold: true
buy_rsi  <- etf_rsi_signals |> filter(signal == "buy")
sell_rsi <- etf_rsi_signals |> filter(signal == "sell")

# Panel 1: price with buy/sell markers
p1 <- highchart(type = "stock") |>
  hc_title(text = "0050.TW — RSI Strategy Signals") |>
  hc_add_series(etf_rsi, type = "line", hcaes(x = date, y = adjusted),
                name = "0050 Price", color = "#234E70") |>
  hc_add_series(buy_rsi,  type = "scatter", hcaes(x = date, y = adjusted),
                name = "Buy",  color = "green", marker = list(symbol = "triangle",      size = 8)) |>
  hc_add_series(sell_rsi, type = "scatter", hcaes(x = date, y = adjusted),
                name = "Sell", color = "red",   marker = list(symbol = "triangle-down", size = 8)) |>
  hc_tooltip(valueDecimals = 2) |>
  hc_legend(enabled = T, align = "center", verticalAlign = "bottom") |>
  hc_add_theme(hc_theme_smpl())

# Panel 2: RSI oscillator with threshold bands
p2 <- highchart() |>
  hc_title(text = "RSI(14) — Overbought / Oversold") |>
  hc_xAxis(type = "datetime") |>
  hc_yAxis(min = 0, max = 100,
           plotLines = list(
             list(value = 70, color = "red",   dashStyle = "ShortDash", width = 1,
                  label = list(text = "Overbought (70)", align = "right")),
             list(value = 30, color = "green", dashStyle = "ShortDash", width = 1,
                  label = list(text = "Oversold (30)",   align = "right"))
           )) |>
  hc_add_series(etf_rsi, type = "line", hcaes(x = date, y = RSI),
                name = "RSI(14)", color = "#8B5CF6") |>
  hc_tooltip(valueDecimals = 2) |>
  hc_add_theme(hc_theme_smpl())

p1
p2
```

------------------------------------------------------------------------

## 4. Bollinger Bands Strategy

**Bollinger Bands** place an envelope around a moving average using the rolling standard deviation of prices. When prices hit the *lower band*, the market is considered locally cheap (buy); when prices hit the *upper band*, it is considered locally expensive (sell).

$$\text{Upper Band} = MA_{20} + 2\,\sigma_{20}, \qquad
\text{Lower Band} = MA_{20} - 2\,\sigma_{20}$$

where $\sigma_{20}$ is the 20-day rolling standard deviation of the adjusted close.

### **4.1. Compute Bollinger Bands**

```{r}
bb <- BBands(etf_data$adjusted, n = 20, sd = 2)

etf_bb <- etf_data |> 
  mutate(
    BB_lower = bb[, "dn"],
    BB_mid   = bb[, "mavg"],
    BB_upper = bb[, "up"],
    pct_b    = bb[, "pctB"]   # percent-B: 0 = lower band, 1 = upper band
  )
head(etf_bb)
```

### **4.2. Generate Bollinger Band Signals**

We buy when price crosses *below* the lower band and sell when price crosses *above* the upper band.

```{r}
etf_bb_signals <- etf_bb |> 
  mutate(
    signal = case_when(
      lag(adjusted) >= lag(BB_lower) & adjusted < BB_lower ~ "buy",
      lag(adjusted) <= lag(BB_upper) & adjusted > BB_upper ~ "sell",
      TRUE ~ NA_character_
    )
  )

etf_bb_signals |> 
  filter(!is.na(signal)) |> 
  select(date, adjusted, BB_lower, BB_upper, signal)
```

### **4.3. Simulate the Bollinger Band Strategy**

```{r}
position <- 0
cash     <- capital
portfolio_bb <- data.frame()

for (i in seq_len(nrow(etf_bb_signals))) {
  date_i  <- etf_bb_signals$date[i]
  price_i <- etf_bb_signals$adjusted[i]
  sig     <- etf_bb_signals$signal[i]

  if (!is.na(sig)) {
    if (sig == "buy") {
      position_new <- floor(cash / price_i)
      position <- position + position_new
      cash     <- cash - position_new * price_i
    } else if (sig == "sell") {
      cash     <- cash + position * price_i
      position <- 0
    }
  }
  total_value <- cash + position * price_i
  portfolio_bb <- bind_rows(portfolio_bb,
                   data.frame(date = date_i, price = price_i,
                              position, cash, total_value))
}

final_bb  <- tail(portfolio_bb$total_value, 1)
return_bb <- (final_bb - capital) / capital * 100
cat("Bollinger Bands Strategy — Final Value: NT$", round(final_bb, 0),
    "| Total Return:", round(return_bb, 2), "%\n")
```

### **4.4. Visualize Bollinger Bands and Signals**

```{r}
#| code-fold: true
buy_bb  <- etf_bb_signals |> filter(signal == "buy")
sell_bb <- etf_bb_signals |> filter(signal == "sell")

highchart(type = "stock") |>
  hc_title(text = "0050.TW — Bollinger Bands Strategy") |>
  hc_add_series(etf_bb, type = "line", hcaes(x = date, y = adjusted),
                name = "0050 Price",   color = "#234E70") |>
  hc_add_series(etf_bb, type = "line", hcaes(x = date, y = BB_upper),
                name = "Upper Band",   color = "#E07A5F", dashStyle = "ShortDash") |>
  hc_add_series(etf_bb, type = "line", hcaes(x = date, y = BB_mid),
                name = "MA20 (Middle)", color = "#9CA3AF", dashStyle = "ShortDot") |>
  hc_add_series(etf_bb, type = "line", hcaes(x = date, y = BB_lower),
                name = "Lower Band",   color = "#10B981", dashStyle = "ShortDash") |>
  hc_add_series(buy_bb,  type = "scatter", hcaes(x = date, y = adjusted),
                name = "Buy",  color = "green", marker = list(symbol = "triangle",      size = 8)) |>
  hc_add_series(sell_bb, type = "scatter", hcaes(x = date, y = adjusted),
                name = "Sell", color = "red",   marker = list(symbol = "triangle-down", size = 8)) |>
  hc_tooltip(valueDecimals = 2, shared = FALSE) |>
  hc_legend(enabled = T, align = "center", verticalAlign = "bottom") |>
  hc_add_theme(hc_theme_smpl())
```

------------------------------------------------------------------------

## 5. Dollar-Cost Averaging (DCA)

The three strategies above all require market-timing decisions. Dollar-Cost Averaging (DCA) abandons timing entirely. Instead, the investor commits a **fixed amount of money** — say NT\$30,000 — on the first trading day of each month, regardless of whether the market is up or down.

The key insight is that the same fixed amount buys *more shares* when prices are low and *fewer shares* when prices are high, automatically creating a form of anti-cyclical investing.

### **5.1. Simulate Monthly DCA**

We set aside the full NT\$1,000,000 as the cash reserve from day one and deploy NT\$30,000 on the first trading day of each month. Any undeployed cash earns no return (a conservative assumption that ignores money-market interest).

```{r}
monthly_invest <- 30000       # NT$ invested each month
total_reserve  <- 1000000     # cash reserve (never more than this deployed in total)

# Identify the first trading day of each month
first_trading_days <- etf_data |>
  mutate(ym = format(date, "%Y-%m")) |>
  group_by(ym) |>
  slice_min(date, n = 1) |>
  ungroup() |>
  select(date) |>
  pull()

# Simulate DCA
cash_dca      <- total_reserve
shares_dca    <- 0
portfolio_dca <- data.frame()
total_invested <- 0   # track cumulative cash deployed

for (i in seq_len(nrow(etf_data))) {
  date_i  <- etf_data$date[i]
  price_i <- etf_data$adjusted[i]

  # On the first trading day of each month, buy as many whole shares as possible
  # with the monthly budget (if cash is still available)
  if (date_i %in% first_trading_days && cash_dca >= monthly_invest) {
    new_shares  <- floor(monthly_invest / price_i)
    cost        <- new_shares * price_i
    shares_dca  <- shares_dca + new_shares
    cash_dca    <- cash_dca   - cost
    total_invested <- total_invested + cost
  }

  total_value <- cash_dca + shares_dca * price_i
  portfolio_dca <- bind_rows(portfolio_dca,
                    data.frame(date = date_i, price = price_i,
                               shares = shares_dca, cash = cash_dca,
                               total_value, total_invested))
}

final_dca  <- tail(portfolio_dca$total_value, 1)
deployed   <- tail(portfolio_dca$total_invested, 1)
return_dca <- (final_dca - capital) / capital * 100
cat("DCA Strategy — Final Value: NT$", round(final_dca, 0),
    "| Total Deployed: NT$", round(deployed, 0),
    "| Total Return:", round(return_dca, 2), "%\n")
```

### **5.2. Shares Accumulated Over Time**

Because DCA buys at varying prices, shares accumulate unevenly. Let us visualize how the position builds month by month.

```{r}
#| code-fold: true
# Monthly summary: date, shares held, price, value
portfolio_dca_monthly <- portfolio_dca |>
  mutate(ym = format(date, "%Y-%m")) |>
  group_by(ym) |>
  slice_max(date, n = 1) |>
  ungroup()

highchart() |>
  hc_title(text = "DCA — Shares Accumulated and Portfolio Value") |>
  hc_xAxis(type = "datetime") |>
  hc_yAxis_multiples(
    list(title = list(text = "Shares Held")),
    list(title = list(text = "Portfolio Value (NT$)"), opposite = TRUE)
  ) |>
  hc_add_series(portfolio_dca_monthly, type = "column",
                hcaes(x = date, y = shares),
                name = "Shares Held", yAxis = 0, color = "#93C5FD") |>
  hc_add_series(portfolio_dca_monthly, type = "line",
                hcaes(x = date, y = total_value),
                name = "Portfolio Value", yAxis = 1, color = "#234E70") |>
  hc_tooltip(shared = TRUE, valueDecimals = 0) |>
  hc_legend(enabled = T, align = "center", verticalAlign = "bottom") |>
  hc_add_theme(hc_theme_smpl())
```

### **5.3. Average Cost Per Share**

A defining feature of DCA is the *average cost per share* — the total cash deployed divided by the total shares accumulated. When prices decline mid-period, DCA buys more shares at cheaper prices, lowering the average cost.

```{r}
avg_cost <- deployed / tail(portfolio_dca$shares, 1)
last_price <- tail(etf_data$adjusted, 1)

cat("Total deployed: NT$", round(deployed, 0), "\n")
cat("Shares accumulated:", tail(portfolio_dca$shares, 1), "\n")
cat("Average cost per share: NT$", round(avg_cost, 2), "\n")
cat("Final price: NT$", round(last_price, 2), "\n")
cat("Unrealized gain per share: NT$", round(last_price - avg_cost, 2), "\n")
```

------------------------------------------------------------------------

## 6. Buy-and-Hold Benchmark

To make the comparison fair, we need a common benchmark. We invest NT\$1,000,000 in 0050.TW on the first available date and hold to the end.

```{r}
initial_bh_price <- etf_data$adjusted[1]
units_bh         <- floor(capital / initial_bh_price)
cash_bh          <- capital - units_bh * initial_bh_price

bh_portfolio <- etf_data |>
  mutate(bh_value = cash_bh + units_bh * adjusted) |>
  select(date, bh_value)

final_bh  <- tail(bh_portfolio$bh_value, 1)
return_bh <- (final_bh - capital) / capital * 100
cat("Buy-and-Hold — Final Value: NT$", round(final_bh, 0),
    "| Total Return:", round(return_bh, 2), "%\n")
```

------------------------------------------------------------------------

## 7. Grand Strategy Comparison

We now bring all five series together — the three technical strategies, DCA, and buy-and-hold — to compare them on equal footing. Because the strategies start at slightly different points (MA requires 50 days of warm-up; RSI requires 14 days; BB requires 20 days), we align all series from the **first date available in the MA strategy** to ensure a common start.

```{r}
start_common <- min(portfolio_ma$date)

# Align all portfolios to the common start date
ma_aligned  <- portfolio_ma |>
  filter(date >= start_common) |>
  select(date, MA_Strategy = total_value)

rsi_aligned <- portfolio_rsi |>
  filter(date >= start_common) |>
  select(date, RSI_Strategy = total_value)

bb_aligned  <- portfolio_bb |>
  filter(date >= start_common) |>
  select(date, BB_Strategy = total_value)

dca_aligned <- portfolio_dca |>
  filter(date >= start_common) |>
  select(date, DCA_Strategy = total_value)

bh_aligned  <- bh_portfolio |>
  filter(date >= start_common) |>
  select(date, BH_0050 = bh_value)

# Merge all
comparison <- ma_aligned |>
  inner_join(rsi_aligned, by = "date") |>
  inner_join(bb_aligned,  by = "date") |>
  inner_join(dca_aligned, by = "date") |>
  inner_join(bh_aligned,  by = "date")

head(comparison)
```

### **7.1. Portfolio Value Comparison**

```{r}
#| code-fold: true
highchart() |>
  hc_title(text = "Strategy Comparison — Portfolio Value (NT$)") |>
  hc_subtitle(text = "All strategies start with NT$1,000,000") |>
  hc_xAxis(type = "datetime") |>
  hc_yAxis(title = list(text = "Portfolio Value (NT$)")) |>
  hc_add_series(comparison, type = "line",
                hcaes(x = date, y = MA_Strategy),
                name = "MA Crossover") |>
  hc_add_series(comparison, type = "line",
                hcaes(x = date, y = RSI_Strategy),
                name = "RSI") |>
  hc_add_series(comparison, type = "line",
                hcaes(x = date, y = BB_Strategy),
                name = "Bollinger Bands") |>
  hc_add_series(comparison, type = "line",
                hcaes(x = date, y = DCA_Strategy),
                name = "DCA") |>
  hc_add_series(comparison, type = "line",
                hcaes(x = date, y = BH_0050),
                name = "Buy-and-Hold") |>
  hc_tooltip(valuePrefix = "NT$", valueDecimals = 0, shared = TRUE) |>
  hc_legend(enabled = T, align = "center", verticalAlign = "bottom") |>
  hc_add_theme(hc_theme_ft())
```

### **7.2. Performance Summary Table**

```{r}
#| code-fold: true
# Annualized volatility: std dev of daily returns * sqrt(252)
daily_ret <- comparison |>
  arrange(date) |>
  mutate(
    r_MA  = MA_Strategy  / lag(MA_Strategy)  - 1,
    r_RSI = RSI_Strategy / lag(RSI_Strategy) - 1,
    r_BB  = BB_Strategy  / lag(BB_Strategy)  - 1,
    r_DCA = DCA_Strategy / lag(DCA_Strategy) - 1,
    r_BH  = BH_0050      / lag(BH_0050)      - 1
  ) |>
  na.omit()

n_days   <- nrow(comparison)
ann_days <- 252

# Mean daily return (annualized) and volatility
summary_stats <- data.frame(
  Strategy       = c("MA Crossover", "RSI", "Bollinger Bands",
                     "DCA", "Buy-and-Hold"),
  Final_Value    = c(final_ma, final_rsi, final_bb, final_dca, final_bh),
  Total_Return   = round(c(return_ma, return_rsi, return_bb, return_dca, return_bh), 2),
  Ann_Vol        = round(c(sd(daily_ret$r_MA),  sd(daily_ret$r_RSI),
                            sd(daily_ret$r_BB),  sd(daily_ret$r_DCA),
                            sd(daily_ret$r_BH)) * sqrt(ann_days) * 100, 2),
  Sharpe         = NA_real_
)

# Sharpe ratio (assume 0% risk-free rate for simplicity)
summary_stats$Sharpe <- round(
  c(mean(daily_ret$r_MA),  mean(daily_ret$r_RSI),
    mean(daily_ret$r_BB),  mean(daily_ret$r_DCA),
    mean(daily_ret$r_BH)) /
  c(sd(daily_ret$r_MA),   sd(daily_ret$r_RSI),
    sd(daily_ret$r_BB),   sd(daily_ret$r_DCA),
    sd(daily_ret$r_BH)) * sqrt(ann_days), 2)

summary_stats$Final_Value <- paste0("NT$", formatC(round(summary_stats$Final_Value),
                                                    format = "d", big.mark = ","))

knitr::kable(summary_stats,
             col.names = c("Strategy", "Final Value", "Total Return (%)",
                           "Ann. Volatility (%)", "Sharpe Ratio"),
             align = "lrrrr",
             caption = "Performance Summary — All Strategies")
```

------------------------------------------------------------------------

## 8. Discussion: What Does This Tell Us?

### 8.1. Timing strategies vs. passive accumulation

The three technical strategies (MA, RSI, Bollinger Bands) all require the investor to be *out of the market* for some stretches. That idle cash earns nothing in our simulation. DCA, by contrast, is always *partially* in the market — it accumulates shares steadily, month by month.

### 8.2. The DCA perspective

DCA is not designed to maximize returns during a bull market — it will lag a lump-sum buy-and-hold when prices trend up monotonically. Its strength lies in **reducing the regret risk** of investing a large lump sum at a market peak. Consider two scenarios:

- **Lump-sum investor** who puts in NT\$1M on January 2, 2023.
- **DCA investor** who puts in NT\$20,000 per month from the same date.

If prices fall mid-year, the DCA investor buys additional shares at the lower price, effectively lowering their average cost. The lump-sum investor is fully exposed from day one and cannot benefit from the dip.

### 8.3. Risk-adjusted performance

The Sharpe ratio is the classic tool for comparing strategies that carry different levels of volatility. A strategy with a lower raw return but much lower volatility can have a *higher* Sharpe ratio than a more volatile winner. Look at the summary table above for the full picture.

### What we ignored

All simulations above share several unrealistic assumptions that a real investor must account for:

- **Transaction costs**: brokerage commissions, securities transaction taxes (0.3% in Taiwan on sales), and market impact.
- **Slippage**: we assume trades execute at the daily close price, which is never guaranteed.
- **Bid-ask spread**: 0050.TW is highly liquid, but the spread still exists.
- **Opportunity cost of idle cash**: in the MA and RSI strategies, cash held between signals earns nothing.

------------------------------------------------------------------------

## Alternative Strategy Designs and Student Exercise

### 💡 Suggested Extensions

1.  **RSI with MA filter**: only act on RSI signals when the price is above MA200 (trending up). This avoids buying into a falling knife.
2.  **MACD**: the `MACD()` function in TTR computes the Moving Average Convergence/Divergence oscillator, which generates its own crossover signals.
3.  **Stochastic Oscillator**: `stoch()` in TTR. Similar logic to RSI but tracks where a closing price sits relative to its recent high-low range.
4.  **DCA with a Boost**: stick to NT\$20,000/month normally, but invest NT\$40,000 whenever RSI \< 30 (doubling down when the market is oversold).
5.  **Value Averaging**: a variant of DCA where the target portfolio value grows by a fixed amount each month; you invest more when the market falls and less (or even sell) when it rises too fast.

### 🧠 Student Exercise

Design and backtest your own strategy for 0050.TW. Your task:

1.  **Define a rule**: use a technical indicator from TTR, a calendar effect, or a DCA variant.
2.  **Implement it**: adapt one of the simulation loops above.
3.  **Benchmark it**: compare to MA crossover, buy-and-hold, and DCA.
4.  **Visualize**: plot all series on one highcharter chart.
5.  **Summarize**: report final value, total return, annualized volatility, and Sharpe ratio.
6.  **Reflect**: in one paragraph, explain *why* your strategy did better or worse, and what you would change next.
