---
title: "Week 4: Portfolio Returns and Diversification"
subtitle: "Financial Calculation with R"
author: "Calvin J. Chiou | 邱健嘉"
institute: "National Chengchi University"
format:
  revealjs:
    theme: serif
    transition: slide
    slide-number: c/t
    progress: true
    footer: "Week 4 · Portfolios"
    chalkboard: false
    code-fold: false
    code-line-numbers: false
    code-overflow: wrap
    highlight-style: github
    embed-resources: true
    logo: images/logo_en.png
    fig-align: center
    width: 1280
    height: 720
knitr:
  opts_chunk:
    echo: true
    # eval: false
    message: false
    warning: false
    fig-align: center
    out-width: "80%"
editor: 
  markdown: 
    wrap: 72
---

```{=html}
<style>
.reveal { font-size: 30px; }
.reveal section h2 { color: #234E70; }
.reveal .footer { font-size: 0.55em; opacity: 0.75; }
.reveal .slide-number { font-size: 0.6em; }
.reveal pre code { font-size: 0.78em; line-height: 1.25; }
.reveal table { font-size: 0.85em; }
.reveal table.dataTable {font-size: 0.75em !important; }
.reveal .dataTables_wrapper {font-size: 0.75em !important; }
</style>
```

```{r, eval=TRUE, echo=FALSE}
#| label: "load"
library(tidyquant)
library(tidyverse)
library(highcharter)
library(ggplot2)
library(ggthemes)
library(DT)
```

# Why Combine Stocks at All?

## The investor's problem

You have NT\$1,000,000 and want to invest it.

You *could* put it all in one stock. But two centuries of theory and
practice say:

**Spread it across many.**

The reason is that combining imperfectly-correlated assets reduces the
*volatility* of your wealth without reducing the expected return
proportionally.

This is the central insight of **modern portfolio theory** (Markowitz,
1952).

# 1. Setting Up a Portfolio

## A four-stock equal-weighted portfolio

```{r}
stocks <- c("XOM", "WMT", "NVDA", "AAPL")
prices <- tq_get(stocks,
                 from = "2023-01-01",
                 to   = "2026-01-01")
```

We have columns:
`symbol, date, open, high, low, close, volume, adjusted`.

```{r, echo=FALSE}
datatable(data = prices, height = "80%", filter = "top",
          options = list(pageLength = 5)) |> 
  formatRound(columns = c("open","high","low","close","adjusted"), digits = 2)
```

## Allocating the budget

With NT\$1,000,000 split equally:

```{r}
budget    <- 1000000
n_stocks  <- length(stocks)
per_stock <- budget / n_stocks    # 250,000 each

initial_positions <- prices |>
  filter(date == min(date)) |>
  mutate(shares = floor(per_stock / adjusted)) |>
  select(symbol, shares)

initial_positions
```

`floor()` because we can only buy whole shares.

# 2. Portfolio Value Over Time

## Tracking total value

```{r}
portfolio <- prices |>
  inner_join(initial_positions, by = "symbol") |>
  mutate(position_value = shares * adjusted) |>
  group_by(date) |>
  summarise(total_value = sum(position_value)) |>
  arrange(date)

head(portfolio)
```

For each date, sum the four positions $\to$ one number per day.

## Portfolio return --- two paths to the same number

### Path A: from total portfolio value

```{r}
portfolio <- portfolio |>
  mutate(R_disc = total_value / lag(total_value) - 1,
         r_log  = log(total_value / lag(total_value)))
```

### Path B: weighted average of individual returns

$$R_p = \sum_{i=1}^{N} w_i \, R_i$$ The cross-section additivity result
from Week 3.

## Path B in code

```{r}
indiv <- prices |>
  group_by(symbol) |>
  mutate(R_disc = adjusted / lag(adjusted) - 1) |>
  ungroup() |>
  mutate(weight = 0.25)   # equal-weighted

port_ret <- indiv |>
  group_by(date) |>
  summarise(R_p = sum(weight * R_disc), .groups = "drop")
```

Path A and Path B will give the same number for discrete returns (up to
rounding from the share floors).

# 3. Portfolio Volatility

## A single stock's volatility

We measure risk by the standard deviation of its returns:

$$\sigma_i = \sqrt{\,\mathrm{Var}(R_i)\,}$$

```{r}
vol_table <- indiv |>
  drop_na(R_disc) |>
  group_by(symbol) |>
  summarise(sigma = sd(R_disc))
vol_table
```

This is the **total** volatility of each stock.

## Portfolio volatility --- the key formula

For an $N$-asset portfolio with weights $\mathbf{w}$ and a covariance
matrix $\boldsymbol{\Sigma}$ of asset returns:

::: {.callout-note title="Portfolio variance and volatility"}
$$\sigma_p^{\,2} = \mathbf{w}^{\top} \boldsymbol{\Sigma}\, \mathbf{w},
\qquad
\sigma_p = \sqrt{\mathbf{w}^{\top} \boldsymbol{\Sigma}\, \mathbf{w}}.$$
:::

**The portfolio's volatility is *not* the weighted average of the
individual volatilities.** The covariances matter.

## The covariance matrix in R

```{r}
returns_wide <- indiv |>
  drop_na(R_disc) |>
  select(symbol, date, R_disc) |>
  pivot_wider(names_from = symbol, values_from = R_disc)

cov_matrix <- cov(returns_wide[ , -1], use = "pairwise.complete.obs")
round(cov_matrix, 6)
```

The diagonal is each stock's variance. The off-diagonals are how the
stocks move *together*.

## Computing $\sigma_p$ via matrix algebra

```{r}
weights <- c(0.25, 0.25, 0.25, 0.25)   # XOM WMT NVDA AAPL

port_var <- t(weights) %*% cov_matrix %*% weights
port_sd  <- sqrt(port_var)
port_sd
```

`%*%` is matrix multiplication.

Compare `port_sd` against `mean(vol_table$sigma)` --- the portfolio is
*less* volatile than the average individual stock. That is
diversification.

# 4. Correlation and Diversification

## Why correlation matters

Decompose the variance formula for $N = 2$:

$$\sigma_p^{\,2}
= w_1^2 \sigma_1^2 + w_2^2 \sigma_2^2
+ 2\, w_1 w_2\, \rho_{12}\, \sigma_1 \sigma_2.$$

- $\rho_{12} = +1$: no diversification,
  $\sigma_p = w_1\sigma_1 + w_2\sigma_2$.
- $\rho_{12} = 0$: substantial reduction.
- $\rho_{12} = -1$: in principle, $\sigma_p$ can be driven to zero.

Real stocks tend to have correlations of 0.3--0.8 with each other ---
enough for diversification to bite, but not enough to eliminate market
risk.

## The correlation matrix

```{r}
cor_matrix <- cor(returns_wide[ , -1], use = "pairwise.complete.obs")
round(cor_matrix, 2)
```

Look at the off-diagonal entries: pairs with lower correlation
contribute more to volatility reduction.

# 5. Comparing Portfolios

## Equal-weighted vs. all-NVDA

```{r}
nvda_only <- prices |>
  filter(symbol == "NVDA", date == min(date)) |>
  mutate(shares = floor(budget / adjusted)) |>
  select(symbol, shares)

nvda_port <- prices |>
  filter(symbol == "NVDA") |>
  inner_join(nvda_only, by = "symbol") |>
  mutate(total_value = shares * adjusted) |>
  select(date, nvda_only = total_value)

combined <- portfolio |>
  select(date, equal_weighted = total_value) |>
  inner_join(nvda_port, by = "date")

head(combined)
```

## Plotting them

```{r}
#| code-fold: true
highchart(width = 1280) |>
  hc_add_series(combined, type = "line",
                hcaes(x = date, y = equal_weighted),
                name = "Equal-Weighted") |>
  hc_add_series(combined, type = "line",
                hcaes(x = date, y = nvda_only),
                name = "100% NVDA") |>
  hc_xAxis(type = "datetime", title = list(text = "Date")) |>
  hc_yAxis(title = list(text = "Portfolio Value (NT$)")) |> 
  hc_tooltip(shared = TRUE, valueDecimals = 0, valuePrefix = "NT$")
```

A single-stock portfolio can crush a diversified one *in a good year*.
It can also collapse spectacularly. Compare *both* return and
volatility, not just return.

## What this week covered

1.  Built an equal-weighted four-stock portfolio with NT\$1M.
2.  Tracked daily portfolio value with `group_by()` + `summarise()`.
3.  Computed portfolio return two ways: from total value, and as
    $\sum w_i R_i$.
4.  Computed portfolio volatility from $\boldsymbol{\Sigma}$:
    $\sigma_p = \sqrt{\mathbf{w}^{\top}\!\boldsymbol{\Sigma}\,\mathbf{w}}$.
5.  Saw how correlation drives diversification gains.
6.  Compared a diversified portfolio against an all-in-one-stock bet.

**Next week:** we ask which part of a stock's volatility is *systematic*
(the market) and which is *idiosyncratic*.

## Exercise

1.  Repeat the four-stock analysis with weights XOM 20%, WMT 30%, NVDA
    30%, AAPL 20%. Recompute portfolio value, return, and $\sigma_p$.
2.  Change the budget to NT\$500,000 with equal weights. Report shares
    purchased per stock.
3.  Add a fifth stock of your choice. Does adding it *raise* or *lower*
    $\sigma_p$? Why?
4.  Plot the four cumulative returns and the portfolio's cumulative
    return on the same `highcharter` chart.

::: {style="text-align:center"}
Thank you!
:::
