---
title: "Week 6: A Working Trading Strategy in R"
subtitle: "Financial Calculation with R"
author: "Calvin J. Chiou | 邱健嘉"
institute: "National Chengchi University"
format:
  revealjs:
    theme: serif
    transition: slide
    slide-number: c/t
    progress: true
    footer: "Week 6 · Trading Strategy"
    chalkboard: false
    code-fold: false
    code-line-numbers: false
    code-overflow: wrap
    highlight-style: github
    embed-resources: true
    logo: images/logo_en.png
    fig-align: center
    width: 1280
    height: 720
knitr:
  opts_chunk:
    echo: true
    # eval: false
    message: false
    warning: false
    fig-align: center
    out-width: "80%"
editor: 
  markdown: 
    wrap: 72
---

```{=html}
<style>
.reveal { font-size: 30px; }
.reveal section h2 { color: #234E70; }
.reveal .footer { font-size: 0.55em; opacity: 0.75; }
.reveal .slide-number { font-size: 0.6em; }
.reveal pre code { font-size: 0.78em; line-height: 1.25; }
.reveal table { font-size: 0.85em; }
.reveal table.dataTable {font-size: 0.75em !important; }
.reveal .dataTables_wrapper {font-size: 0.75em !important; }
</style>
```

```{r, eval=TRUE, echo=FALSE}
#| label: "load"
library(tidyquant)
library(tidyverse)
library(highcharter)
library(ggplot2)
library(DT)
```

# The Capstone

## Where we are

In five weeks we have built:

- R data-handling fluency (Week 1)
- Time value of money (Week 2)
- Return calculations (Week 3)
- Portfolio formation and volatility (Week 4)
- The market model and $\beta$ (Week 5)

This week we combine all of it into one complete pipeline: download data
$\to$ generate signals $\to$ simulate trades $\to$ benchmark.

## The case study

We will trade the Taiwan 50 ETF, ticker `0050.TW` --- one of the most
actively traded ETFs in Taiwan, representing large-cap Taiwanese stocks.

- **Period:** 2023-01-01 to 2026-01-01.
- **Strategy:** a moving-average crossover.
- **Initial capital:** NT\$1,000,000.
- **Benchmarks:** TAIEX index and buy-and-hold 0050.

The goal is not to *get rich* --- it is to learn the *methodology* of
designing, implementing, and honestly evaluating a quantitative
strategy.

# 1. Downloading the Data

## ETF prices

```{r}
etf <- tq_get("0050.TW",
              from = "2023-01-01",
              to   = "2026-01-01")
```

Two columns matter for what follows: `date` and `adjusted`.

```{r, echo=FALSE}
datatable(data = etf, height = "80%",
          options = list(pageLength = 5)) |> 
  formatRound(columns = c("open","high","low","close","adjusted"), digits = 2)
```

# 2. Moving Averages

## What is a moving average?

A rolling mean of the last $n$ days of adjusted close.

::: {.callout-note title="Simple moving average"}
$$MA_n(t) = \frac{1}{n}\sum_{i=0}^{n-1} P_{t-i}$$
:::

We use two:

- **MA20:** short-term trend (\~1 trading month).
- **MA50:** medium-term trend (\~2.5 trading months).

The relative position of the two is our trading signal.

## Computing both in R

```{r}
etf_ma <- etf |>
  mutate(MA20 = SMA(adjusted, n = 20),
         MA50 = SMA(adjusted, n = 50)) |>
  na.omit()
```

`SMA()` comes from the **TTR** package, which `tidyquant` loads for us.

`na.omit()` drops the first 49 rows where MA50 is not yet defined.

```{r, echo=FALSE}
datatable(data = etf_ma, height = "80%",
          options = list(pageLength = 5)) |> 
  formatRound(columns = c("open","high","low","close","adjusted","MA20","MA50"), digits = 2)
```

# 3. Trading Signals

## The crossover rule

::: {.callout-note title="Signal definition"}
- **Buy** when MA20 crosses *above* MA50 ("golden cross").
- **Sell** when MA20 crosses *below* MA50 ("death cross").
:::

In code, "crosses above today" means: *yesterday's MA20 was below MA50,
today's MA20 is at or above MA50.*

## Implementing the rule

```{r}
etf_signals <- etf_ma |>
  mutate(signal = case_when(
    lag(MA20) <  lag(MA50) & MA20 >= MA50 ~ "buy",
    lag(MA20) >  lag(MA50) & MA20 <= MA50 ~ "sell",
    TRUE                                  ~ NA_character_
  ))
```

`case_when()` is the tidyverse multi-branch `if/else`.

```{r, echo=FALSE}
datatable(data = etf_signals |> filter(!is.na(signal)), height = "80%",
          options = list(pageLength = 5)) |> 
  formatRound(columns = c("open","high","low","close","adjusted","MA20","MA50"), digits = 2)
```

# 4. Simulating the Trades

## The all-in / all-out simulator

A simple state machine: at each date, check the signal and update
`cash`, `position`, and `total_value`.

```{r}
capital   <- 1000000
cash      <- capital
position  <- 0
portfolio <- data.frame()

for (i in seq_len(nrow(etf_signals))) {
  date_i  <- etf_signals$date[i]
  price_i <- etf_signals$adjusted[i]
  sig     <- etf_signals$signal[i]

  if (!is.na(sig)) {
    if (sig == "buy") {
      position <- floor(cash / price_i)
      cash     <- cash - position * price_i
    } else if (sig == "sell") {
      cash     <- cash + position * price_i
      position <- 0
    }
  }
  total_value <- cash + position * price_i
  portfolio   <- bind_rows(portfolio,
                  data.frame(date = date_i, price = price_i,
                             position, cash, total_value))
}
```

## What the loop is doing

1.  Walk through every trading day in order.
2.  On a buy signal: spend all the cash on as many shares as we can
    afford.
3.  On a sell signal: liquidate completely back to cash.
4.  Every day, record `total_value = cash + position * price`.

This is the simplest possible "in or out" strategy --- no leverage, no
shorting, no partial positions, no transaction costs.

# 5. Plotting the Strategy

## Portfolio value over time

```{r}
#| code-fold: true
highchart(width = 1280) |>
  hc_title(text = "MA Crossover --- Portfolio Value") |>
  hc_xAxis(type = "datetime",
           title = list(text = "Date")) |>
  hc_yAxis(title = list(text = "Total Value (NT$)")) |>
  hc_add_series(data = portfolio, type = "line",
                hcaes(x = date, y = total_value),
                name = "Portfolio Value") |>
  hc_tooltip(valuePrefix = "NT$", valueDecimals = 2) |>
  hc_legend(enabled = FALSE)
```

## Final performance number

```{r}
final_value <- tail(portfolio$total_value, 1)
return_pct  <- (final_value - capital) / capital * 100

cat("Final value: NT$", round(final_value, 2), "\n")
cat("Total return:", round(return_pct, 2), "%\n")
```

One number is never enough --- we need a *benchmark*.

# 6. Benchmarking

## The TAIEX market benchmark

```{r}
market <- tq_get("^TWII",
                 from = "2023-01-01",
                 to   = "2026-01-01") |>
  select(date, market_price = adjusted)

init_price <- market$market_price[1]
units      <- floor(1000000 / init_price)

market_bh <- market |>
  mutate(benchmark_value = units * market_price) |>
  select(date, benchmark_value)
```

A simulated NT\$1M invested in the index, never traded.

## Buy-and-hold 0050

```{r}
bh_etf <- etf |>
  filter(date >= ymd("2023-03-27"))

init_etf <- bh_etf$adjusted[1]
units_bh <- floor(1000000 / init_etf)

bh_etf <- bh_etf |>
  mutate(bh_etf_value = units_bh * adjusted) |>
  select(date, bh_etf_value)
```

We start all three strategies on the same date so the comparison is
fair.

## Merging and plotting

```{r}
comp <- portfolio |>
  select(date, strategy_value = total_value) |>
  inner_join(market_bh,  by = "date") |>
  inner_join(bh_etf,     by = "date")
```

```{r}
#| code-fold: true
highchart(type = "stock", width = 1280) |>
  hc_title(text = "Strategy vs. Benchmarks") |>
  hc_xAxis(type = "datetime") |>
  hc_yAxis(title = list(text = "Portfolio Value (NT$)")) |>
  hc_add_series(comp, "line",
                hcaes(x = date, y = strategy_value),
                name = "MA Strategy") |>
  hc_add_series(comp, "line",
                hcaes(x = date, y = bh_etf_value),
                name = "Buy-and-Hold 0050") |>
  hc_add_series(comp, "line",
                hcaes(x = date, y = benchmark_value),
                name = "TAIEX") |>
  hc_tooltip(
    valuePrefix = "NT$",
    valueDecimals = 0,  # 2-decimal tooltips
    shared = TRUE   # show all series at once
  )
```

# 7. Lessons and Limitations

## What this kind of comparison teaches

Three serious lessons emerge from looking at the chart honestly:

- The MA crossover often *lags* sharp trend reversals --- it buys late
  and sells late.
- A long stretch of sideways prices generates many false signals
  ("whipsaws"), each costing implicit slippage.
- Buy-and-hold, the most boring strategy on the planet, is competitive
  with --- sometimes better than --- an active rule.

This is the *standard* result for naive technical strategies in academic
backtests. It is a feature of the problem, not a bug in your code.

## What we ignored

- **Transaction costs and slippage**: every trade costs money in
  reality.
- **Bid-ask spreads**: you don't actually trade at the close.
- **Survivorship bias**: 0050 is the post-2003 winner index.
- **Look-ahead bias**: we used `lag()` correctly --- always check this.
- **Risk adjustment**: we compared raw returns, not Sharpe ratios. A
  volatile winner is not the same kind of "good" as a smooth winner.

A serious research replication would address every one of these.

## Five strategy variants to try

1.  **Add a stop-loss:** exit if the position drops 10% from entry.
2.  **Momentum allocation:** switch between 0050 and a bond proxy based
    on trailing 12-month return.
3.  **Volatility-scaled positions:** larger position when realized
    volatility is low; smaller when high.
4.  **Mean reversion:** use Bollinger Bands or RSI to fade extremes
    instead of follow trends.
5.  **Calendar rotation:** invest only Nov--Apr; cash the rest.

## What this week --- and this course --- covered

1.  Built a complete strategy: download $\to$ signal $\to$ trade $\to$
    benchmark.
2.  Used `SMA()` and `case_when()` for signal generation.
3.  Walked a `for` loop through every trading day to update positions.
4.  Compared the strategy fairly against two benchmarks.
5.  Discussed honest caveats about backtests.

You have now built, end-to-end, the workflow that every quantitative
finance team in the world starts with. The difference is in the polish
--- not the bones.

## Final exercise: your own strategy

**Design and backtest your own rule.**

1.  Choose a signal --- another MA pair, a momentum window, a calendar
    effect, an RSI threshold, or anything else.
2.  Modify the simulation loop from this deck.
3.  Benchmark against the MA crossover strategy, buy-and-hold 0050, and
    TAIEX.
4.  Plot all four series on one `highcharter` chart.
5.  Report final value, total return, and one risk number (annualized
    $\sigma$ or Sharpe).
6.  In one paragraph, explain *why* your strategy did better or worse,
    and what you would change next.

::: {style="text-align:center"}
Thank you, and happy investing!
:::
