---
title: "Week 2: The Time Value of Money"
subtitle: "Financial Calculation with R"
author: "Calvin J. Chiou | 邱健嘉"
institute: "National Chengchi University"
format:
  revealjs:
    theme: serif
    transition: slide
    slide-number: c/t
    progress: true
    footer: "Week 2 · Time Value of Money"
    chalkboard: false
    code-fold: false
    code-line-numbers: false
    code-overflow: wrap
    highlight-style: github
    embed-resources: true
    logo: images/logo_en.png
    fig-align: center
    # fig-width: 8
    # fig-height: 4.5
    width: 1280
    height: 720
knitr:
  opts_chunk:
    echo: true
    # eval: false
    message: false
    warning: false
    fig-align: center
    out-width: "80%"
editor: 
  markdown: 
    wrap: 72
---

```{=html}
<style>
.reveal { font-size: 30px; }
.reveal section h2 { color: #234E70; }
.reveal .footer { font-size: 0.55em; opacity: 0.75; }
.reveal .slide-number { font-size: 0.6em; }
.reveal pre code { font-size: 0.78em; line-height: 1.25; }
.reveal table { font-size: 0.85em; }
.reveal table.dataTable {font-size: 0.75em !important; }
.reveal .dataTables_wrapper {font-size: 0.75em !important; }
</style>
```

```{r, eval=TRUE, echo=FALSE}
#| label: "load"
library(tidyquant)
library(tidyverse)
library(highcharter)
library(ggplot2)
library(ggthemes)
```

# The Core Idea

## A dollar today is not a dollar tomorrow

Suppose someone offers you a choice:

- **Option A:** NT\$1,000 today.
- **Option B:** NT\$1,000 one year from today.

Which is better? Most people pick A. Why?

## A dollar today is not a dollar tomorrow

Suppose someone offers you a choice:

- **Option A:** NT\$1,000 today.
- **Option B:** NT\$1,000 one year from today.

Which is better? Most people pick A. Why?

**Three reasons:**

1.  You could invest A and earn interest.
2.  Inflation may erode B's purchasing power.
3.  B is uncertain --- you might not actually get it.

## The fundamental relation

If you can invest at an annual rate $r$, then **NT\$1 today** grows to
**NT\$(1+r)** in one year.

Equivalently, **NT\$1 next year** is worth only **NT\$**$\frac{1}{1+r}$
today.

This single idea --- the **discount factor** $\frac{1}{1+r}$ --- is the
foundation of every valuation formula in finance, from a savings account
to a billion-dollar acquisition.

# 1. Future Value and Present Value

## Future Value

Investing $PV$ today at rate $r$ for $n$ years:

::: {.callout-note title="Future value (annual compounding)"}
$$FV = PV \cdot (1+r)^n$$
:::

```{r}
PV <- 1000
r  <- 0.05
n  <- 10
FV <- PV * (1 + r)^n
FV
```

NT\$1,000 today becomes NT\$1,628.89 in ten years at 5%.

## Present Value

Running the relation backwards:

::: {.callout-note title="Present value"}
$$PV = \frac{FV}{(1+r)^n}$$
:::

```{r}
FV <- 1000
r  <- 0.05
n  <- 10
PV <- FV / (1 + r)^n
PV
```

NT\$1,000 *promised* in ten years is worth only NT\$613.91 today.

## Writing a reusable function

Hard-coding numbers is fragile. Wrap the formula in a function:

```{r}
pv <- function(FV, r, n) FV / (1 + r)^n
fv <- function(PV, r, n) PV * (1 + r)^n

pv(FV = 1000, r = 0.05, n = 10)
fv(PV = 1000, r = 0.05, n = 10)
```

Now any calculation in the course can call `pv()` or `fv()`.

# 2. Compounding

## Compounding $m$ times per year

If interest is paid $m$ times per year, each period the rate is $r/m$:

::: callout-note
$$FV = PV \cdot \left(1 + \frac{r}{m}\right)^{m\,n}$$
:::

```{r}
PV <- 1000; r <- 0.05; n <- 10
m <- c(1, 2, 4, 12, 365)
FV_m <- PV * (1 + r/m)^(m*n)
data.frame(periods_per_year = m, FV = round(FV_m, 2))
```

More frequent compounding $\Rightarrow$ larger FV, but the gains shrink
fast.

## Continuous compounding

Take $m \to \infty$:

::: {.callout-note title="Continuous compounding"}
$$FV = PV \cdot e^{r n}, \qquad PV = FV \cdot e^{-r n}$$
:::

```{r}
PV <- 1000; r <- 0.05; n <- 10
FV_cont <- PV * exp(r * n)
FV_cont
```

This is the form you'll meet again in option pricing (Black-Scholes uses
continuous compounding throughout).

## Visualizing the compounding gap

```{r}
#| code-fold: true
n <- 0:30
df <- data.frame(
  year   = n,
  simple = 1000 + 1000 * 0.05 * n,
  comp   = 1000 * (1 + 0.05)^n,
  comp_monthly = 1000 * (1 + 0.05/12)^(n*12),
  cont   = 1000 * exp(0.05 * n)
)
```

```{r, eval=TRUE}
#| code-fold: true
highchart(width = 1280) |>
  hc_title(text = "Simple vs. Annual vs. Continuous Compounding") |>
  hc_xAxis(title = list(text = "Year"), categories = df$year) |>
  hc_yAxis(title = list(text = "Future Value (NT$)")) |>
  hc_add_series(name = "Simple", data = df$simple, type = "line") |>
  hc_add_series(name = "Annual",data = df$comp,type = "line") |>
  hc_add_series(name = "Monthly",data = df$comp_monthly,type = "line") |>
  hc_add_series(name = "Continuous", data = df$cont, type = "line") |>
  hc_tooltip(shared = TRUE, valueDecimals = 2) |>
  hc_plotOptions(series = list(marker = list(enabled = FALSE)))
```

# 3. Annuities

## What is an annuity?

An **annuity** is a stream of equal cash flows $C$ paid at the end of
each period for $n$ periods. Examples:

- Monthly mortgage payments.
- Yearly retirement-fund withdrawals.
- Coupon payments on a bond.

The present value of all those payments is the sum of $n$ discounted
cash flows.

## The annuity formula

::: {.callout-note title="Present value of an ordinary annuity"}
$$PV = C \cdot \frac{1 - (1+r)^{-n}}{r}$$
:::

```{r}
annuity_pv <- function(C, r, n) {
  C * (1 - (1 + r)^(-n)) / r
}
annuity_pv(C = 50000, r = 0.04, n = 20)
```

NT\$50,000 every year for 20 years at 4% has a present value of about
NT\$679,500 --- much less than $20 \times 50{,}000 = 1{,}000{,}000$.

## Perpetuity

Let $n \to \infty$ in the annuity formula:

::: {.callout-note title="Perpetuity"}
$$PV = \frac{C}{r}$$
:::

```{r}
perp_pv <- function(C, r) C / r
perp_pv(C = 50000, r = 0.04)
```

A NT\$50,000 perpetual payment at 4% is worth NT\$1,250,000 today ---
the same logic underlies the **Gordon growth model** for valuing stocks.

# 4. Net Present Value

## Mixed cash flows: NPV

Most real projects produce *unequal* cash flows at different times. Sum
them all, discounted to today:

::: {.callout-note title="Net present value"}
$$NPV = \sum_{t=0}^{n} \frac{CF_t}{(1+r)^t}$$
:::

**Decision rule:** invest if $NPV > 0$.

## NPV in R

```{r}
cf <- c(-100000, 30000, 35000, 40000, 45000)
r  <- 0.08
t  <- 0:(length(cf) - 1)
NPV <- sum(cf / (1 + r)^t)
NPV
```

A small investment with a positive NPV of about NT\$22,300 --- the
project creates value.

## Internal Rate of Return (IRR)

The IRR is the discount rate that makes $NPV = 0$:
$$\sum_{t=0}^{n} \frac{CF_t}{(1+\text{IRR})^t} = 0.$$

```{r}
npv <- function(r, cf) {
  t <- 0:(length(cf) - 1)
  sum(cf / (1 + r)^t)
}
irr <- uniroot(npv, interval = c(-0.99, 1), cf = cf)$root
irr
```

If IRR exceeds your required return, accept the project --- equivalent
to $NPV > 0$ when cash flows are conventional.

# 5. A Mini Application

## How much to save for retirement?

You are 25, want NT\$10,000,000 by age 65, earning 6% per year. How much
do you need to save *each year*?

Rearrange the future-value-of-annuity formula:

$$FV = C \cdot \frac{(1+r)^n - 1}{r}
\quad\Rightarrow\quad
C = \frac{FV \cdot r}{(1+r)^n - 1}$$

```{r}
target <- 10000000; r <- 0.06; n <- 40
C <- target * r / ((1 + r)^n - 1)
C
```

Only about NT\$64,600 per year --- the power of compounding over 40
years.

## What if you wait until 35?

```{r}
target <- 10000000; r <- 0.06; n <- 30
C <- target * r / ((1 + r)^n - 1)
C
```

NT\$126,500 per year --- nearly *double*, just by starting 10 years
later.

**Lesson:** time is the most expensive variable in any financial plan.

## The monthly payment formula

A fully-amortizing loan of principal $L$ at monthly rate $r_m = r/12$
over $N$ months has constant monthly payment:

$$M = L \cdot \frac{r_m \cdot (1+r_m)^N}{(1+r_m)^N - 1}$$

This is just the annuity-payment formula from earlier, applied monthly.

```{r}
monthly_payment <- function(L, annual_rate, years) {
  r_m <- annual_rate / 12
  N   <- years * 12
  L * r_m * (1 + r_m)^N / ((1 + r_m)^N - 1)
}
```

For a **3-year grace period**, the first 36 months are interest-only
(monthly payment = $L \cdot r_m$), and the remaining principal then
amortizes over 27 years --- a second call to `monthly_payment()` with
`years = 27`.

## **Comparing the six scenarios**

```{r}
#| code-fold: true
loans <- c(20e6, 15e6, 10e6)
rate  <- 0.022

# All six (loan × scheme) combinations
scenarios <- tibble(
  loan   = rep(loans, each = 2),
  scheme = rep(c("No grace", "3-yr grace"), times = 3)
) |>
  mutate(
    pmt_yr1to3 = if_else(scheme == "No grace",
                         monthly_payment(loan, rate, 30),
                         loan * rate / 12),                          # interest only
    pmt_yr4on  = if_else(scheme == "No grace",
                         monthly_payment(loan, rate, 30),
                         monthly_payment(loan, rate, 27)),           # amortize over 27 yr
    total_paid     = pmt_yr1to3 * 36 + pmt_yr4on * (30 * 12 - 36),
    total_interest = total_paid - loan
  )
```

```{r, echo=FALSE}
library(DT)
scenarios |> 
  datatable(
    filter = "none", height = "80%"
  ) |> 
  formatRound(columns = c(3,4,5,6), digits = 2)
```

The grace period **lowers the first three years**, but the post-grace
payment is **higher** (the same principal must amortize in 27 years
instead of 30) --- and **total interest paid is larger**.

## Where does the money actually go?

Each monthly payment splits between *interest* and *principal*. In the
early years almost all of it is interest --- so the debt balance shrinks
very slowly. The closed-form remaining balance after $k$ months is
$$B_k = L \cdot (1+r_m)^k - M \cdot \frac{(1+r_m)^k - 1}{r_m}$$

```{r}
# Remaining balance after k months for a fully-amortizing loan
balance_amort <- function(L, annual_rate, years, k) {
  r_m <- annual_rate / 12
  M   <- monthly_payment(L, annual_rate, years)
  L * (1 + r_m)^k - M * ((1 + r_m)^k - 1) / r_m
}
# remaining after 24 years (24*12 months)
balance_amort(20e6, 0.025, 30, 24*12)
# remaining after 10 years
balance_amort(20e6, 0.025, 30, 10*12)
```

## Remaining principal over 30 years {.smaller}

```{r}
#| code-fold: true
# Full 360-month trajectory for one (loan, scheme) pair
balance_path <- function(L, annual_rate, scheme) {
  months <- 0:(30 * 12)
  if (scheme == "No grace") {
    bal <- balance_amort(L, annual_rate, 30, months)
  } else {
    bal <- ifelse(months <= 36, L,
                  balance_amort(L, annual_rate, 27, months - 36))
  }
  data.frame(month = months, balance = bal, loan = L, scheme = scheme)
}
balances <- do.call(rbind, lapply(loans, function(L) {
  rbind(balance_path(L, rate, "No grace"),
        balance_path(L, rate, "3-yr grace"))
}))
```

```{r}
#| code-fold: true
#| fig-height: 5
# Named color lookup: one color per loan size (millions of NT$)
colors_by_loan <- c("20" = "#1F3A5F", "15" = "#2E86AB", "10" = "#E07A5F")
# Start an empty highcharter chart with all the chart-level styling
hc <- highchart(width = 1280) |>
  hc_title(text = "Debt balance (remaining principal) over 30 years") |>
  hc_subtitle(text = "Dashed = 3-year grace period; solid = no grace") |>
  hc_xAxis(title = list(text = "Month")) |>
  hc_yAxis(title = list(text = "Balance (NT$ million)"),
           labels = list(format = "{value:.1f}")) |>
  hc_tooltip(shared = TRUE, valueDecimals = 2,
             valuePrefix = "NT$", valueSuffix = " M") |>
  hc_plotOptions(line = list(marker = list(enabled = FALSE)))

# Add one series per (loan size, scheme) pair --- six series in total.
# Same color for the same loan size; solid line = no grace, dashed = 3-yr grace.
for (L in loans) {
  for (sch in c("No grace", "3-yr grace")) {
    d <- balances |> dplyr::filter(loan == L, scheme == sch)
    hc <- hc |> hc_add_series(
      data      = d,
      hcaes(x = month, y = balance / 1e6),
      name      = paste0("NT$", L / 1e6, "M · ", sch),
      type      = "line",
      color     = colors_by_loan[as.character(L / 1e6)],
      dashStyle = if (sch == "No grace") "Solid" else "ShortDash"
    )
  }
}

# Print the finished chart
hc
```

The grace-period borrower carries the full principal for 3 years, then
has to pay it down faster. The dashed line is flat from month 0 to month
36, then descends more steeply than its solid counterpart.

## Visualizing the payment paths

```{r}
#| code-fold: true
#| fig-width: 10
#| fig-height: 4.5
paths <- expand.grid(loan   = loans,
                     scheme = c("No grace", "3-yr grace"),
                     month  = 1:(30 * 12)) |>
  mutate(
    payment = case_when(
      scheme == "No grace"                 ~ monthly_payment(loan, rate, 30),
      scheme == "3-yr grace" & month <= 36 ~ loan * rate / 12,
      scheme == "3-yr grace" & month >  36 ~ monthly_payment(loan, rate, 27)
    ),
    loan_label = paste0("NT$", loan / 1e6, "M")
  )

# Compute the average monthly payment per (loan, scheme) over the full 30 years
payment_summary <- paths |>
  group_by(loan_label, scheme) |>
  summarise(avg_payment = mean(payment), .groups = "drop")

ggplot(payment_summary,
       aes(x = loan_label, y = avg_payment / 1000, fill = scheme)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = scales::comma(round(avg_payment / 1000))),
            position = position_dodge(width = 0.7),
            vjust = -0.4, size = 3.5) +
  scale_fill_manual(values = c("No grace"   = "#234E70",
                               "3-yr grace" = "#D9534F")) +
  labs(title = "Average monthly mortgage payment over 30 years",
       subtitle = "By loan size and grace-period scheme",
       x = "Loan size", y = "Average monthly payment (NT$ thousand)",
       fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

## The takeaway

The grace period is **not free money** --- it is a *re-timing* of cash
flows. You pay less in years 1--3 and more in years 4--30, plus extra
interest over the full life of the loan.

**When does a grace period make sense?**

- You expect your income to rise significantly within a few years.
- You face large one-time expenses early in the loan (renovation, kids).
- You can invest the cash saved early at a *higher* rate than 2.2% ---
  the same time-value-of-money logic in reverse.

**When does it not?**

- You can already afford the level payment comfortably.
- You want to minimize total interest paid over the life of the loan.

The Week 4 lesson on **investment opportunity cost** will sharpen this
intuition further.

## What this week covered

1.  Why money has a time value: $\frac{1}{1+r}$ is the discount factor.
2.  PV, FV, and the effect of compounding frequency.
3.  Continuous compounding: $FV = PV\,e^{rn}$.
4.  Annuities, perpetuities, and the Gordon-growth intuition.
5.  NPV and IRR for evaluating mixed cash-flow projects.
6.  Built a mortgage calculator and compared loan structures.
7.  Wrote reusable functions: `pv()`, `fv()`, `annuity_pv()`,
    `monthly_payment()`, `npv()`.

**Next week:** we apply this thinking to *stock returns*.

## Exercise

1.  You inherit NT\$500,000 today. Invested at 7%, what is it worth in
    30 years?
2.  A bond pays NT\$10,000 a year for 15 years, then NT\$1,000,000 at
    maturity. What is its present value at $r = 5\%$?
3.  Write a function `annuity_fv(C, r, n)`. Verify it against your
    `fv()` and `annuity_pv()` functions.
4.  A project costs NT\$2 million today and returns NT\$600,000 a year
    for 5 years. Is it worth accepting at $r = 9\%$? What is the IRR?

::: {style="text-align:center"}
Thank you!
:::
