---
title: "Week 3: Calculating Returns"
subtitle: "Financial Calculation with R"
author: "Calvin J. Chiou | 邱健嘉"
institute: "National Chengchi University"
format:
  revealjs:
    theme: serif
    transition: slide
    slide-number: c/t
    progress: true
    footer: "Week 3 · Returns"
    chalkboard: false
    code-fold: false
    code-line-numbers: false
    code-overflow: wrap
    highlight-style: github
    embed-resources: true
    logo: images/logo_en.png
    fig-align: center
    # fig-width: 8
    # fig-height: 4.5
    width: 1280
    height: 720
knitr:
  opts_chunk:
    echo: true
    # eval: false
    message: false
    warning: false
    fig-align: center
    out-width: "80%"
editor: 
  markdown: 
    wrap: 72
---

```{=html}
<style>
.reveal { font-size: 30px; }
.reveal section h2 { color: #234E70; }
.reveal .footer { font-size: 0.55em; opacity: 0.75; }
.reveal .slide-number { font-size: 0.6em; }
.reveal pre code { font-size: 0.78em; line-height: 1.25; }
.reveal table { font-size: 0.85em; }
.reveal table.dataTable {font-size: 0.75em !important; }
.reveal .dataTables_wrapper {font-size: 0.75em !important; }
</style>
```

```{r, eval=TRUE, echo=FALSE}
#| label: "load"
library(tidyquant)
library(tidyverse)
library(highcharter)
library(ggplot2)
library(ggthemes)
library(DT)
```

# What Is a Return?

## The investor's most basic question

You bought a stock at NT\$100. It's now NT\$110.

**How much did you make?**

Two equally valid answers:

- In money: NT\$10.
- In percentage: 10%.

This week is about *percentage returns* --- and the fact that there are
**two different definitions**, both correct, both used in practice.

# 1. Discrete (Simple) Returns

## Definition

::: {.callout-note title="Discrete return"}
$$R_t = \frac{P_t}{P_{t-1}} - 1 = \frac{P_t - P_{t-1}}{P_{t-1}}$$
:::

The intuitive percentage change. If $P_{t-1} = 100$ and $P_t = 110$,
then $R_t = 10\%$.

This is the version you see in newspapers, brokerage statements, and
industry research.

## Computing it in R with `dplyr`

```{r}
apple <- tq_get("AAPL", from = "2023-01-01", to = "2026-01-01")

apple <- apple |>
  arrange(date) |>
  mutate(R_disc = adjusted / lag(adjusted) - 1)
```

`lag()` shifts the column down one row --- the previous day's price.

```{r, echo=FALSE}
datatable(data = apple |> select(date, adjusted, R_disc), 
          height = "80%",
          options = list(pageLength = 5)) |> 
  formatRound(columns = c("adjusted","R_disc"), digits = 4)
```

# 2. Continuous (Log) Returns

## Definition

::: {.callout-note title="Continuous (log) return"}
$$r_t = \ln\!\left(\frac{P_t}{P_{t-1}}\right) = \ln P_t - \ln P_{t-1}$$
:::

The log return is the *instantaneous-growth-rate* version of return ---
the counterpart of continuous compounding from Week 2.

For small changes, $r_t \approx R_t$. They diverge for large moves.

## Discrete vs. continuous --- side by side

```{r}
apple <- apple |>
  mutate(R_disc = adjusted / lag(adjusted) - 1,
         r_log  = log(adjusted / lag(adjusted)))
```

For a typical daily move of $\pm 1\%$, the two are nearly identical. For
a $+50\%$ jump, $R = 0.50$ but $r = \ln(1.5) = 0.405$.

```{r, echo=FALSE}
datatable(data = apple |> select(date, R_disc, r_log), 
          height = "80%",
          options = list(pageLength = 5)) |> 
  formatRound(columns = c("R_disc","r_log"), digits = 4)
```

## When to use which?

|                   | **Discrete** $R$    | **Continuous** $r$           |
|-------------------|---------------------|------------------------------|
| Interpretation    | Simple % change     | Log-based, continuous        |
| Time-additive?    | **No**              | **Yes**                      |
| Cross-section sum | **Yes** (portfolio) | No                           |
| Where used        | Industry, reporting | Academic research, modelling |

The *time-additivity* of log returns is what makes them the standard in
econometrics: a 2-day log return is just the sum of two daily log
returns.

# 3. Time vs. Cross-section Additivity

## Why log returns are time-additive

Over $n$ days: $$\frac{P_T}{P_0} = \prod_{t=1}^{T}\frac{P_t}{P_{t-1}}
\quad\Rightarrow\quad
\ln\!\frac{P_T}{P_0} = \sum_{t=1}^{T} r_t.$$ So
$r_{0 \to T} = r_1 + r_2 + \cdots + r_T$. **Sum them.**

But the *same* formula in discrete form is:
$$1 + R_{0 \to T} = \prod_{t=1}^{T} (1 + R_t).$$ So
$R_{0 \to T} \ne R_1 + R_2 + \cdots + R_T$. **Compound them.**

## Why discrete returns are cross-section-additive

For a portfolio with weights $w_i$ summing to 1:

$$R_p = \sum_{i=1}^{N} w_i\,R_i \qquad \text{(exactly)}$$ This is exact
for discrete returns, and only approximate for log returns.

**Practical rule of thumb:**

- Aggregating *across stocks* at one point in time $\to$ discrete.
- Aggregating *across time* for one stock $\to$ continuous.

# 4. Multi-period Returns

## Cumulative discrete return

$$R_{\text{cum}} = \prod_{t=1}^{T}(1 + R_t) - 1$$

```{r}
apple <- apple |>
  mutate(cum_R = cumprod(1 + coalesce(R_disc, 0)) - 1)
```

`cumprod()` does the running product; `coalesce(R_disc, 0)` replaces the
first NA with 0 so the cumulative product starts from 1.

## Cumulative log return

$$r_{\text{cum}} = \sum_{t=1}^{T} r_t$$

```{r}
apple <- apple |>
  mutate(cum_r = cumsum(coalesce(r_log, 0)))
```

Just `cumsum()` --- nothing more.

To convert back to a discrete cumulative return:
$R_{\text{cum}} = e^{r_{\text{cum}}} - 1$.

# 5. Buy-and-Hold Return

## A passive strategy in one line

The simplest possible "strategy": buy on day 0, never trade. The return
over any horizon is just the price ratio:

$$R_{\text{BH}} = \frac{P_T}{P_0} - 1$$

```{r}
apple |>
  summarise(BH_return = last(adjusted) / first(adjusted) - 1)
```

It sounds trivial --- but as we'll see in Week 6, a well-chosen
buy-and-hold often beats more complex active strategies. Don't dismiss
it.

# 6. Visualizing Returns

## A return time-series

```{r}
#| code-fold: true
highchart(type = "stock", width = 1280) |>
  hc_add_series(apple, type = "line",
                hcaes(x = date, y = R_disc * 100),
                name = "AAPL Daily Return (%)") |>
  hc_title(text = "AAPL Daily Returns 2023") |>
  hc_yAxis(title = list(text = "Return (%)")) |> 
  hc_tooltip(valueDecimals = 2, valueSuffix = " %")
```

## Cumulative-return chart

```{r}
#| code-fold: true
highchart(type = "stock", width = 1280) |>
  hc_add_series(apple, type = "line",
                hcaes(x = date, y = cum_R * 100),
                name = "Cumulative Discrete Return") |>
  hc_add_series(apple, type = "line",
                hcaes(x = date, y = cum_r * 100),
                name = "Cumulative Log Return") |>
  hc_title(text = "AAPL 2023 --- Two Ways to Measure Cumulative Return") |>
  hc_yAxis(title = list(text = "%")) |> 
  hc_tooltip(valueDecimals = 2, valueSuffix = " %")
```

The two lines diverge slightly as returns accumulate --- the visual
proof of compounding.

## What this week covered

1.  Two return definitions: discrete $R$ and continuous $r$.
2.  Time-additivity (logs sum) vs. cross-section additivity (discrete
    sums).
3.  Multi-period returns: `cumprod()` and `cumsum()`.
4.  Buy-and-hold return in one line.
5.  Plotting return series with `highcharter`.

**Next week:** we put many stocks into a single portfolio and compute
its return *and* its volatility.

## Exercise

1.  Download MSFT, GOOGL, and NVDA for 2023. Compute both discrete and
    log daily returns for each.
2.  For NVDA, compute the cumulative 2023 return two ways (`cumprod` of
    discrete vs. `cumsum` of log). By how much do they differ at
    year-end?
3.  What is the *annualized* mean daily log return for each stock?
    (Hint: multiply daily mean by 252.)
4.  Plot the three cumulative-return paths on one `highcharter` chart.
    Which stock won 2023?

::: {style="text-align:center"}
Thank you!
:::
