---
title: "Week 5: Risk, Beta, and the Market Model"
subtitle: "Financial Calculation with R"
author: "Calvin J. Chiou | 邱健嘉"
institute: "National Chengchi University"
format:
  revealjs:
    theme: serif
    transition: slide
    slide-number: c/t
    progress: true
    footer: "Week 5 · Risk & Beta"
    chalkboard: false
    code-fold: false
    code-line-numbers: false
    code-overflow: wrap
    highlight-style: github
    embed-resources: true
    logo: images/logo_en.png
    fig-align: center
    width: 1280
    height: 720
knitr:
  opts_chunk:
    echo: true
    # eval: false
    message: false
    warning: false
    fig-align: center
    out-width: "80%"
editor: 
  markdown: 
    wrap: 72
---

```{=html}
<style>
.reveal { font-size: 30px; }
.reveal section h2 { color: #234E70; }
.reveal .footer { font-size: 0.55em; opacity: 0.75; }
.reveal .slide-number { font-size: 0.6em; }
.reveal pre code { font-size: 0.78em; line-height: 1.25; }
.reveal table { font-size: 0.85em; }
.reveal table.dataTable {font-size: 0.75em !important; }
.reveal .dataTables_wrapper {font-size: 0.75em !important; }
</style>
```

```{r, eval=TRUE, echo=FALSE}
#| label: "load"
library(tidyquant)
library(tidyverse)
library(highcharter)
library(ggplot2)
library(ggthemes)
library(DT)
```

# Two Kinds of Risk

## Last week we measured volatility. This week we *decompose* it.

A stock's total volatility $\sigma_i$ has two ingredients:

- **Systematic risk:** co-moves with the market. Cannot be diversified
  away.
- **Idiosyncratic risk:** stock-specific. Vanishes in a well-diversified
  portfolio.

The **market model** is the simplest tool to separate the two. And it
gives us a single number that captures systematic risk: $\beta$.

# 1. The Market Model

## A linear regression of stock on market

::: {.callout-note title="Market model"}
$$R_{i,t} = \alpha_i + \beta_i\, R_{m,t} + \varepsilon_{i,t}$$
:::

- $R_{i,t}$: stock $i$'s return on day $t$.
- $R_{m,t}$: the market's return on day $t$.
- $\alpha_i$: intercept --- average performance unrelated to the market.
- $\beta_i$: slope --- the sensitivity of the stock to the market.
- $\varepsilon_{i,t}$: residual --- the part the market cannot explain.

In R, this is a one-line call to `lm()`.

## What does $\beta$ mean?

$$\beta_i = \frac{\mathrm{Cov}(R_i,\,R_m)}{\mathrm{Var}(R_m)}$$

- $\beta > 1$: stock amplifies market moves --- more cyclical.
- $\beta = 1$: stock moves with the market on average.
- $0 < \beta < 1$: stock moves with the market but less than 1-for-1 ---
  defensive.
- $\beta < 0$: stock moves *against* the market --- rare; gold and some
  hedges.

For tech stocks like NVDA we typically see $\beta > 1$; for utilities
and consumer staples, $\beta < 1$.

# 2. Estimating $\beta$ in R

## The data

```{r}
stock_ticker  <- "NVDA"
market_ticker <- "^GSPC"   # S&P 500
start_date    <- "2020-01-01"
end_date      <- "2026-01-01"

stock_data  <- tq_get(stock_ticker,  from = start_date, to = end_date)
market_data <- tq_get(market_ticker, from = start_date, to = end_date)
```

Three years of daily data is a typical estimation window.

## Computing daily returns

```{r}
stock_returns <- stock_data |>
  tq_transmute(select     = adjusted,
               mutate_fun = periodReturn,
               period     = "daily",
               col_rename = "stock_return")

market_returns <- market_data |>
  tq_transmute(select     = adjusted,
               mutate_fun = periodReturn,
               period     = "daily",
               col_rename = "market_return")

returns <- left_join(stock_returns, market_returns,
                     by = "date") |>
  na.omit()
```

`tq_transmute()` with `periodReturn` is the tidyquant shorthand for the
discrete-return formula from Week 3.

## Running the regression

```{r}
model <- lm(stock_return ~ market_return, data = returns)
summary(model)
```

What to look for in the output:

- **Estimate** for `market_return`: this is $\hat\beta$.
- **Estimate** for `(Intercept)`: this is $\hat\alpha$.
- **Pr(\>\|t\|)**: is $\beta$ statistically distinguishable from 0?
- **Multiple R-squared**: fraction of variance explained by the market.

## Extracting $\hat\beta$ cleanly

```{r}
library(broom)

beta_hat <- tidy(model) |>
  filter(term == "market_return") |>
  pull(estimate)

cat("Estimated beta for", stock_ticker, "=",
    round(beta_hat, 2), "\n")
```

For NVDA over 2020--2022, this comes out around 1.6 --- NVDA tended to
move about 1.6% for every 1% move in the S&P 500.

# 3. Visualizing the Regression

## A scatter plot with the fitted line

```{r}
#| code-fold: true
returns$fit <- predict(model)

highchart(width = 1280) |>
  hc_title(text = "NVDA vs. S&P 500 --- Linear Fit") |>
  hc_xAxis(title = list(text = "Market Return")) |>
  hc_yAxis(title = list(text = "Stock Return")) |>
  hc_add_series(data = returns, type = "scatter",
                hcaes(x = market_return, y = stock_return),
                name = "Daily Returns") |>
  hc_add_series(data = returns[order(returns$market_return), ],
                type = "line",
                hcaes(x = market_return, y = fit),
                name = "Fitted line") |> 
  hc_tooltip(valueDecimals = 2)
```

The *slope* of that line is your eyeball estimate of $\beta$.

# 4. Idiosyncratic Risk

## What the regression *can't* explain

The residuals
$\hat\varepsilon_{i,t} = R_{i,t} - \hat{\alpha} - \hat\beta\, R_{m,t}$
are the stock-specific portion of returns.

::: {.callout-note title="Idiosyncratic volatility"}
$$\sigma_{\varepsilon,i} = \mathrm{sd}\!\left(\hat\varepsilon_{i,t}\right)$$
:::

```{r}
residuals_i <- resid(model)
idio_vol    <- sd(residuals_i)
cat("Idiosyncratic volatility:",
    round(idio_vol, 4), "\n")
```

## Decomposition of variance

In population:

$$\underbrace{\sigma_i^2}_{\text{total}}
\;=\;
\underbrace{\beta_i^2\, \sigma_m^2}_{\text{systematic}}
\;+\;
\underbrace{\sigma_{\varepsilon,i}^2}_{\text{idiosyncratic}}.$$

```{r}
sigma_total <- sd(returns$stock_return)
sigma_mkt   <- sd(returns$market_return)
sys_part    <- beta_hat^2 * sigma_mkt^2
idio_part   <- idio_vol^2

c(total_var = sigma_total^2,
  systematic = sys_part,
  idiosyncratic = idio_part)
```

`systematic + idiosyncratic` should approximately equal `total_var` ---
the analysis-of-variance identity behind the market model.

# 5. Why $\beta$ Matters

## $\beta$ is what the CAPM prices

The Capital Asset Pricing Model (Sharpe, 1964) says expected excess
return depends only on $\beta$:

::: {.callout-note title="CAPM"}
$$E[R_i] - R_f = \beta_i \,\bigl(E[R_m] - R_f\bigr)$$
:::

- High-$\beta$ stocks should earn higher expected returns.
- Idiosyncratic risk should *not* be priced --- you can diversify it
  away.

The empirical evidence is mixed (the "low-beta anomaly"), but $\beta$
remains the single most-cited measure of a stock's market risk.

## Course connections

What we built this week will be the foundation for two later ideas
you'll meet if you continue:

- **Cumulative abnormal returns (CAR)** in event studies. The market
  model predicts a stock's "normal" return; whatever the stock does
  *above and beyond* that prediction around an event date is the
  *abnormal return*.
- **Multi-factor models** (Fama-French, Carhart). We replace the single
  market return with several factor portfolios. The R syntax barely
  changes: `lm(R_i ~ MKT + SMB + HML)`.

## What this week covered

1.  Decomposed total risk into systematic and idiosyncratic parts.
2.  Estimated $\beta$ with one call to `lm()`.
3.  Extracted and interpreted `summary(model)` and `tidy(model)`.
4.  Computed idiosyncratic volatility from residuals.
5.  Connected $\beta$ to the CAPM and to multi-factor extensions.

**Next week:** we use everything we've learned to design, run, and
benchmark a trading strategy.

## Exercise

1.  Estimate $\beta$ for AAPL, KO (Coca-Cola), and TSLA against the S&P
    500 over 2020--2022. Rank them. Which is the most market-sensitive?
2.  Estimate $\beta$ for 2330.TW (TSMC) against `^TWII` (TAIEX) over the
    same window.
3.  For each stock, report $\beta$, $\sigma_{\varepsilon}$, and the
    $R^2$ of the regression.
4.  Plot the four scatter plots side-by-side and comment on which
    stock's returns are best explained by its respective market index.

::: {style="text-align:center"}
Thank you!
:::
